---
title: "PTvsLM"
execute: 
  message: false
  warning: false
format: 
  html:
    code-fold: true
    
---

```{r Loading packages}
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
library(hdf5r)
library(SeuratData)
library(ggrepel)
library(tidyverse)
library(decontX)
library(SingleCellExperiment)
library(pheatmap)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(clusterProfiler)
library(enrichplot)
library(DOSE)
library(enrichplot)
library(MAST)
library(lme4)
library(DoMultiBarHeatmap)
library(RColorBrewer)
library(colorRamp2)
library(scCustomize)
library(DoubletFinder)
library(dendextend)
library(ggcorrplot)
library(ggtext)
library(grid)
library(ComplexHeatmap)
library(DESeq2)
library(muscat)
library(harmony)
library(cowplot)
```
## Importing and orginising data

```{r Loading data, DecontX, adding metaData}
#Setting working directory
setwd("~/T10")

#Importing data from files
#LM files
filtLM1 <- Read10X_h5(filename ="data/LM/PDX1/1_JEPDX1_6/sample_filtered_feature_bc_matrix.h5")
filtLM2 <- Read10X_h5(filename ="data/LM/PDX1/2_CHPDX1_5/sample_filtered_feature_bc_matrix.h5")
filtLM4 <- Read10X_h5(filename ="data/LM/PDX2/4_JEPDX2_5/sample_filtered_feature_bc_matrix.h5")
filtLM5 <- Read10X_h5(filename ="data/LM/PDX2/5_CHPDX2_5/sample_filtered_feature_bc_matrix.h5")
filtLM6 <- Read10X_h5(filename ="data/LM/PDX2/6_CHPDX2_7/sample_filtered_feature_bc_matrix.h5")
filtLM8 <- Read10X_h5(filename ="data/LM/PDX3/8_JEPDX3_5/sample_filtered_feature_bc_matrix.h5")
filtLM9 <- Read10X_h5(filename ="data/LM/PDX3/9_CHPDX3_3/sample_filtered_feature_bc_matrix.h5")
filtLM10 <- Read10X_h5(filename ="data/LM/PDX3/10_CHPDX3_5/sample_filtered_feature_bc_matrix.h5")

#PT files
filtPT3A <- Read10X("data/PT/PDX1/3_JEPDX1_7/#A (17th February)/filtered_feature_bc_matrix/")
filtPT3B <- Read10X("data/PT/PDX1/3_JEPDX1_7/#B (7th March)/filtered_feature_bc_matrix/")
filtPT4 <- Read10X("data/PT/PDX2/4_JEPDX2_5/filtered_feature_bc_matrix/")
filtPT5 <- Read10X_h5(filename ="data/PT/PDX2/5_CHPDX2_5/filtered_feature_bc_matrix.h5")
filtPT7 <- Read10X_h5(filename ="data/PT/PDX2/7_CHPDX2_2/filtered_feature_bc_matrix.h5")
filtPT8 <- Read10X_h5(filename ="data/PT/PDX3/8_JEPDX3_5/filtered_feature_bc_matrix.h5")
filtPT9 <- Read10X_h5(filename ="data/PT/PDX3/9_CHPDX3_3/filtered_feature_bc_matrix.h5")
filtPT10 <- Read10X_h5(filename ="data/PT/PDX3/10_CHPDX3_5/filtered_feature_bc_matrix.h5")

#Creating SeuratObjects
serLM1 <- CreateSeuratObject(filtLM1, project = "1_JEPDX1_6")
serLM2 <- CreateSeuratObject(filtLM2, project = "2_CHPDX1_5")
serLM4 <- CreateSeuratObject(filtLM4, project = "4_JEPDX2_5")
serLM5 <- CreateSeuratObject(filtLM5, project = "5_CHPDX2_5")
serLM6 <- CreateSeuratObject(filtLM6, project = "6_CHPDX2_7")
serLM8 <- CreateSeuratObject(filtLM8, project = "8_JEPDX3_5")
serLM9 <- CreateSeuratObject(filtLM9, project = "9_CHPDX3_3")
serLM10 <- CreateSeuratObject(filtLM10, project = "10_CHPDX3_5")
serPT3A <- CreateSeuratObject(filtPT3A, project = "3_JEPDX1_7A")
serPT3B <- CreateSeuratObject(filtPT3B, project = "3_JEPDX1_7B")
serPT4 <- CreateSeuratObject(filtPT4, project = "4_JEPDX2_5")
serPT5 <- CreateSeuratObject(filtPT5, project = "5_CHPDX2_5")
serPT7 <- CreateSeuratObject(filtPT7, project = "7_CHPDX2_2")
serPT8 <- CreateSeuratObject(filtPT8, project = "8_JEPDX3_5")
serPT9 <- CreateSeuratObject(filtPT9, project = "9_CHPDX3_3")
serPT10 <- CreateSeuratObject(filtPT10, project = "10_CHPDX3_5")


#Importing data from raw files
#raw LM files
rawLM1 <- Read10X_h5(filename ="data/LM/PDX1/1_JEPDX1_6/sample_raw_feature_bc_matrix.h5")
rawLM2 <- Read10X_h5(filename ="data/LM/PDX1/2_CHPDX1_5/sample_raw_feature_bc_matrix.h5")
rawLM4 <- Read10X_h5(filename ="data/LM/PDX2/4_JEPDX2_5/sample_raw_feature_bc_matrix.h5")
rawLM5 <- Read10X_h5(filename ="data/LM/PDX2/5_CHPDX2_5/sample_raw_feature_bc_matrix.h5")
rawLM6 <- Read10X_h5(filename ="data/LM/PDX2/6_CHPDX2_7/sample_raw_feature_bc_matrix.h5")
rawLM8 <- Read10X_h5(filename ="data/LM/PDX3/8_JEPDX3_5/sample_raw_feature_bc_matrix.h5")
rawLM9 <- Read10X_h5(filename ="data/LM/PDX3/9_CHPDX3_3/sample_raw_feature_bc_matrix.h5")
rawLM10 <- Read10X_h5(filename ="data/LM/PDX3/10_CHPDX3_5/sample_raw_feature_bc_matrix.h5")

#raw PT files
rawPT3A <- Read10X("data/PT/PDX1/3_JEPDX1_7/#A (17th February)/raw_feature_bc_matrix/")
rawPT3B <- Read10X("data/PT/PDX1/3_JEPDX1_7/#B (7th March)/raw_feature_bc_matrix/")
rawPT4 <- Read10X("data/PT/PDX2/4_JEPDX2_5/raw_feature_bc_matrix/")
rawPT5 <- Read10X_h5(filename ="data/PT/PDX2/5_CHPDX2_5/raw_feature_bc_matrix.h5")
rawPT7 <- Read10X_h5(filename ="data/PT/PDX2/7_CHPDX2_2/raw_feature_bc_matrix.h5")
rawPT8 <- Read10X_h5(filename ="data/PT/PDX3/8_JEPDX3_5/raw_feature_bc_matrix.h5")
rawPT9 <- Read10X_h5(filename ="data/PT/PDX3/9_CHPDX3_3/raw_feature_bc_matrix.h5")
rawPT10 <- Read10X_h5(filename ="data/PT/PDX3/10_CHPDX3_5/raw_feature_bc_matrix.h5")

#skapar SingleCellExperiment för att kunna köra DecontX, ta bort ambient RNA
sceLM1.filt <- SingleCellExperiment(list(counts = filtLM1))
sceLM2.filt <- SingleCellExperiment(list(counts = filtLM2))
sceLM4.filt <- SingleCellExperiment(list(counts = filtLM4))
sceLM5.filt <- SingleCellExperiment(list(counts = filtLM5))
sceLM6.filt <- SingleCellExperiment(list(counts = filtLM6))
sceLM8.filt <- SingleCellExperiment(list(counts = filtLM8))
sceLM9.filt <- SingleCellExperiment(list(counts = filtLM9))
sceLM10.filt <- SingleCellExperiment(list(counts = filtLM10))
sceLM1.raw <- SingleCellExperiment(list(counts = rawLM1))
sceLM2.raw <- SingleCellExperiment(list(counts = rawLM2))
sceLM4.raw <- SingleCellExperiment(list(counts = rawLM4))
sceLM5.raw <- SingleCellExperiment(list(counts = rawLM5))
sceLM6.raw <- SingleCellExperiment(list(counts = rawLM6))
sceLM8.raw <- SingleCellExperiment(list(counts = rawLM8))
sceLM9.raw <- SingleCellExperiment(list(counts = rawLM9))
sceLM10.raw <- SingleCellExperiment(list(counts = rawLM10))
#samma för PT
scePT3A.filt <- SingleCellExperiment(list(counts = filtPT3A))
scePT3B.filt <- SingleCellExperiment(list(counts = filtPT3B))
scePT4.filt <- SingleCellExperiment(list(counts = filtPT4))
scePT5.filt <- SingleCellExperiment(list(counts = filtPT5))
scePT7.filt <- SingleCellExperiment(list(counts = filtPT7))
scePT8.filt <- SingleCellExperiment(list(counts = filtPT8))
scePT9.filt <- SingleCellExperiment(list(counts = filtPT9))
scePT10.filt <- SingleCellExperiment(list(counts = filtPT10))
scePT3A.raw <- SingleCellExperiment(list(counts = rawPT3A))
scePT3B.raw <- SingleCellExperiment(list(counts = rawPT3B))
scePT4.raw <- SingleCellExperiment(list(counts = rawPT4))
scePT5.raw <- SingleCellExperiment(list(counts = rawPT5))
scePT7.raw <- SingleCellExperiment(list(counts = rawPT7))
scePT8.raw <- SingleCellExperiment(list(counts = rawPT8))
scePT9.raw <- SingleCellExperiment(list(counts = rawPT9))
scePT10.raw <- SingleCellExperiment(list(counts = rawPT10))

rm(filtLM1, filtLM2, filtLM4, filtLM5, filtLM6, filtLM8, filtLM9, filtLM10, filtPT3A, filtPT3B, filtPT4, filtPT5, filtPT7, filtPT8, filtPT9, filtPT10, rawLM1, rawLM2, rawLM4, rawLM5, rawLM6, rawLM8, rawLM9, rawLM10, rawPT3A, rawPT3B, rawPT4, rawPT5, rawPT7, rawPT8, rawPT9, rawPT10)

#Kör DecontX med raw_matrix som bakgrund
sceLM1.result <- decontX(sceLM1.filt, background = sceLM1.raw)
sceLM2.result <- decontX(sceLM2.filt, background = sceLM2.raw)
sceLM4.result <- decontX(sceLM4.filt, background = sceLM4.raw)
sceLM5.result <- decontX(sceLM5.filt, background = sceLM5.raw)
sceLM6.result <- decontX(sceLM6.filt, background = sceLM6.raw)
sceLM8.result <- decontX(sceLM8.filt, background = sceLM8.raw)
sceLM9.result <- decontX(sceLM9.filt, background = sceLM9.raw)
sceLM10.result <- decontX(sceLM10.filt, background = sceLM10.raw)
scePT3A.result <- decontX(scePT3A.filt, background = scePT3A.raw)
scePT3B.result <- decontX(scePT3B.filt, background = scePT3B.raw)
scePT4.result <- decontX(scePT4.filt, background = scePT4.raw)
scePT5.result <- decontX(scePT5.filt, background = scePT5.raw)
scePT7.result <- decontX(scePT7.filt, background = scePT7.raw)
scePT8.result <- decontX(scePT8.filt, background = scePT8.raw)
scePT9.result <- decontX(scePT9.filt, background = scePT9.raw)
scePT10.result <- decontX(scePT10.filt, background = scePT10.raw)

options(Seurat.object.assay.calcn = TRUE)

#Adding decontX corrected into the count matrix of the SeuratObjects
serLM1[["RNA"]]$counts <- round(decontXcounts(sceLM1.result))
serLM2[["RNA"]]$counts <- round(decontXcounts(sceLM2.result))
serLM4[["RNA"]]$counts <- round(decontXcounts(sceLM4.result))
serLM5[["RNA"]]$counts <- round(decontXcounts(sceLM5.result))
serLM6[["RNA"]]$counts <- round(decontXcounts(sceLM6.result))
serLM8[["RNA"]]$counts <- round(decontXcounts(sceLM8.result))
serLM9[["RNA"]]$counts <- round(decontXcounts(sceLM9.result))
serLM10[["RNA"]]$counts <- round(decontXcounts(sceLM10.result))

serPT3A[["RNA"]]$counts <- round(decontXcounts(scePT3A.result))
serPT3B[["RNA"]]$counts <- round(decontXcounts(scePT3B.result))
serPT4[["RNA"]]$counts <- round(decontXcounts(scePT4.result))
serPT5[["RNA"]]$counts <- round(decontXcounts(scePT5.result))
serPT7[["RNA"]]$counts <- round(decontXcounts(scePT7.result))
serPT8[["RNA"]]$counts <- round(decontXcounts(scePT8.result))
serPT9[["RNA"]]$counts <- round(decontXcounts(scePT9.result))
serPT10[["RNA"]]$counts <- round(decontXcounts(scePT10.result))

rm(sceLM1.filt, sceLM1.raw, sceLM1.result, sceLM2.filt, sceLM2.raw, sceLM2.result, sceLM4.filt, sceLM4.raw, sceLM4.result, sceLM5.filt, sceLM5.raw, sceLM5.result, sceLM6.filt, sceLM6.raw, sceLM6.result, sceLM8.filt, sceLM8.raw, sceLM8.result, sceLM9.filt, sceLM9.raw, sceLM9.result, sceLM10.filt, sceLM10.raw, sceLM10.result, scePT3A.filt, scePT3A.raw, scePT3A.result, scePT3B.filt, scePT3B.raw, scePT3B.result, scePT4.filt, scePT4.raw, scePT4.result, scePT5.filt, scePT5.raw, scePT5.result, scePT7.filt, scePT7.raw, scePT7.result, scePT8.filt, scePT8.raw, scePT8.result, scePT9.filt, scePT9.raw, scePT9.result, scePT10.filt, scePT10.raw, scePT10.result)
gc()

#Addning information layers
serLM1$PDX <- "LU-NB-1"
serLM2$PDX <- "LU-NB-1"
serLM4$PDX <- "LU-NB-2"
serLM5$PDX <- "LU-NB-2"
serLM6$PDX <- "LU-NB-2"
serLM8$PDX <- "LU-NB-3"
serLM9$PDX <- "LU-NB-3"
serLM10$PDX <- "LU-NB-3"
serPT3A$PDX <- "LU-NB-1"
serPT3B$PDX <- "LU-NB-1"
serPT4$PDX <- "LU-NB-2"
serPT5$PDX <- "LU-NB-2"
serPT7$PDX <- "LU-NB-2"
serPT8$PDX <- "LU-NB-3"
serPT9$PDX <- "LU-NB-3"
serPT10$PDX <- "LU-NB-3"

serLM1$type <- "LM"
serLM2$type <- "LM"
serLM4$type <- "LM"
serLM5$type <- "LM"
serLM6$type <- "LM"
serLM8$type <- "LM"
serLM9$type <- "LM"
serLM10$type <- "LM"
serPT3A$type <- "PT"
serPT3B$type <- "PT"
serPT4$type <- "PT"
serPT5$type <- "PT"
serPT7$type <- "PT"
serPT8$type <- "PT"
serPT9$type <- "PT"
serPT10$type <- "PT"

serLM1$mouse <- "1"
serLM2$mouse <- "2"
serLM4$mouse <- "4"
serLM5$mouse <- "5"
serLM6$mouse <- "6"
serLM8$mouse <- "8"
serLM9$mouse <- "9"
serLM10$mouse <- "10"
serPT3A$mouse <- "3"
serPT3B$mouse <- "3"
serPT4$mouse <- "4"
serPT5$mouse <- "5"
serPT7$mouse <- "7"
serPT8$mouse <- "8"
serPT9$mouse <- "9"
serPT10$mouse <- "10"

serLM1$batch <- "B1"
serLM2$batch <- "B1"
serLM4$batch <- "B1"
serLM5$batch <- "B1"
serLM6$batch <- "B2"
serLM8$batch <- "B2"
serLM9$batch <- "B2"
serLM10$batch <- "B2"
serPT3A$batch <- "B3"
serPT3B$batch <- "B4"
serPT4$batch <- "B5"
serPT5$batch <- "B6"
serPT7$batch <- "B7"
serPT8$batch <- "B8"
serPT9$batch <- "B9"
serPT10$batch <- "B10"

serLM1$PDX_type <- "LM LU-NB-1"
serLM2$PDX_type <- "LM LU-NB-1"
serLM4$PDX_type <- "LM LU-NB-2"
serLM5$PDX_type <- "LM LU-NB-2"
serLM6$PDX_type <- "LM LU-NB-2"
serLM8$PDX_type <- "LM LU-NB-3"
serLM9$PDX_type <- "LM LU-NB-3"
serLM10$PDX_type <- "LM LU-NB-3"
serPT3A$PDX_type <- "PT LU-NB-1"
serPT3B$PDX_type <- "PT LU-NB-1"
serPT4$PDX_type <- "PT LU-NB-2"
serPT5$PDX_type <- "PT LU-NB-2"
serPT7$PDX_type <- "PT LU-NB-2"
serPT8$PDX_type <- "PT LU-NB-3"
serPT9$PDX_type <- "PT LU-NB-3"
serPT10$PDX_type <- "PT LU-NB-3"

serLM1$type_mouse <- "LM1"
serLM2$type_mouse <- "LM2"
serLM4$type_mouse <- "LM4"
serLM5$type_mouse <- "LM5"
serLM6$type_mouse <- "LM6"
serLM8$type_mouse <- "LM8"
serLM9$type_mouse <- "LM9"
serLM10$type_mouse <- "LM10"
serPT3A$type_mouse <- "PT3A"
serPT3B$type_mouse <- "PT3B"
serPT4$type_mouse <- "PT4"
serPT5$type_mouse <- "PT5"
serPT7$type_mouse <- "PT7"
serPT8$type_mouse <- "PT8"
serPT9$type_mouse <- "PT9"
serPT10$type_mouse <- "PT10"

#Merging into 1 object
serAll <- merge(serLM1, c(serLM2, serLM4, serLM5, serLM6, serLM8, serLM9, serLM10, serPT3A, serPT3B, serPT4, serPT5, serPT7, serPT8, serPT9, serPT10), add.cell.ids = c("LM1", "LM2", "LM4", "LM5", "LM6", "LM8", "LM9", "LM10", "PT3A", "PT3B", "PT4", "PT5", "PT7", "PT8", "PT9", "PT10"))
serAll <- JoinLayers(serAll)

#object with only PT to be able to compensate for chemistry difference and intron inclusion as newer run include more genes
serPT <- merge(serPT3A, c(serPT3B, serPT4, serPT7, serPT8), add.cell.ids = c("PT3A", "PT3B", "PT4", "PT7", "PT8"))
serPT <- JoinLayers(serPT)

#object with only LM to be able to compensate for chemistry difference and intron inclusion as newer run include more genes
serLM <- merge(serLM1, c(serLM2, serLM4, serLM5, serLM6, serLM8, serLM9, serLM10), add.cell.ids = c("LM1", "LM2", "LM4", "LM5", "LM6", "LM8", "LM9", "LM10"))
serLM <- JoinLayers(serLM)

#Remove unecessary objects
rm(serLM1, serLM2, serLM4, serLM5, serLM6, serLM8, serLM9, serLM10, serPT3A, serPT3B, serPT4, serPT5, serPT7, serPT8, serPT9, serPT10)
gc()

#Showing size of dataset
dim(serAll)
table(serAll$type_mouse)
table(serAll$PDX_type)
table(serAll$type)

```
Size of original dataset
###########################################################################################################
##Filtering
```{r Checking size}
length(rownames(serAll))
length(rownames(serPT))
length(rownames(serLM))
```
Skillnaden i antal gener analyserade hos PT och LM. Delvis beroende på chemistry

```{r Filtering after decontX}
#genesLM <- rownames(serLM)
#genesPT <- rownames(serPT)
#sharedGenes = intersect(genesLM, genesPT)
#filteredData <- subset(serAll, features = sharedGenes) #Keeping only genes expressed both in PT and in LM in some degree. 

countMatrix <- GetAssayData(serAll, assay = "RNA", layer = "counts")
countMatrix[is.na(countMatrix)] <- 0
filteredData <- SetAssayData(serAll, assay = "RNA", layer = "counts", new.data = countMatrix) #Remove NA's introduced in decontX. 

filteredData <- subset(filteredData, nFeature_RNA > 0) #Remove cells with no expressed genes (the expression was removed in decontX as cells were decontamination)
#Showing size of dataset
dim(filteredData)
```
Antal gener och celler efter exklusion av gener endast i LM och endast i PT, chemistry v4. 

```{r Calculate and plot features}
rm(serPT, sharedGenes, countMatrix)
gc()

#Showing Mitochondrial, Ribosomal, Hemoglobin, Platelate markers
filteredData <- PercentageFeatureSet(filteredData, pattern = "^MT-", col.name = "percent_mito")
filteredData <- PercentageFeatureSet(filteredData, "^RP[SL]", col.name = "percent_ribo")
filteredData <- PercentageFeatureSet(filteredData, "^HB[^(P|E|S)]", col.name = "percent_hb")
filteredData <- PercentageFeatureSet(filteredData, "PECAM1|PF4", col.name = "percent_plat")

#Creating violin plots of QC variables
feats <- c("nFeature_RNA", "nCount_RNA", "percent_mito", "percent_ribo", "percent_hb", "percent_plat")
NormalizeData(filteredData)
VlnPlot(filteredData, group.by = "type_mouse", split.by = "type", features = feats, pt.size = 0.05, ncol = 3)

```

We see high mitochondrial gene expression in many cells, this is indicative of cell membrane damage, allowing free RNA to leave the cytoplasm but not mitochondias due to their size. We want to filter out these cells. As well as cells with low total expression and low ribosomal expression. 

We can also limit the amount of genes we analyse by excluding genes expressed by only a few cells, as they will not hold any significant information for analysis. We will also exclude mitochondrial, ribosomal, and hemoglobin genes as physiological fluctuations in these may influence our analysis as well. 

```{r Filter based on features}
#Filter out cells and genes with low expression
selectCell <- WhichCells(filteredData, expression = nFeature_RNA > 100)
selectFeat <- rownames(filteredData)[Matrix::rowSums(filteredData[["RNA"]]$counts) > 10]
filteredData <- subset(filteredData, features = selectFeat, cells = selectCell)

#Filter out cells with high mitocondrial and low ribosomal expression
filteredData <- subset(filteredData, percent_mito < 20 & percent_ribo > 5)
```
Filtering done:
*Remove cells with less than 100 genes expressed.
*Remove genes expressed by less than 10 cells.
*Remove cells with above 20% of gene expression being mitochondrial genes.
*Remove cells with less than 5% of gene expression being ribosomal genes.

```{r Plotting filtered data}
#Showing size of dataset
dim(filteredData)
table(filteredData$type_mouse)
table(filteredData$PDX_type)
table(filteredData$type)


#New violin plot of QC variables
feats <- c("nFeature_RNA", "nCount_RNA", "percent_mito", "percent_ribo", "percent_hb")
VlnPlot(filteredData, group.by = "type_mouse", features = feats, pt.size = 0.05, ncol = 3) + NoLegend()
```
Size of dataset and visualization of quality control variables after first filtering. 

```{r Filter out genes}
# Filter Mitochondrial, Ribosomal and Hemoglobin genes
filteredData <- filteredData[!grepl("^MT-", rownames(filteredData)), ]
filteredData <- filteredData[!grepl("^RP[SL]", rownames(filteredData)), ]
filteredData <- filteredData[!grepl("^HB[^(P|E|S)]", rownames(filteredData)), ]
dim(filteredData)
```

Filtering done:
*Remove mitochondrial genes
*Remove ribosomal genes
*Remove Hemoglobin genes
Showing new dimensions fo dataset.

```{r Calculate doublets (DoubletFinder)}

run_doubletfinder_custom <- function(seu_sample_subset, multiplet_rate = NULL){
  # for debug
  #seu_sample_subset <- samp_split[[1]]
  # Print sample number
  print(paste0("Sample ", unique(seu_sample_subset[['batch']]), '...........')) 
  
  #Manually setting predicted doublet rate based on recovered cells per batch
  if(unique(seu_sample_subset[['batch']]) == 'B1'){
    multiplet_rate <- 0.008
  }
  if(unique(seu_sample_subset[['batch']]) == 'B2'){
    multiplet_rate <- 0.008
  }
  if(unique(seu_sample_subset[['batch']]) == 'B3'){
    multiplet_rate <- 0.023
  }
  if(unique(seu_sample_subset[['batch']]) == 'B4'){
    multiplet_rate <- 0.008
  }
  if(unique(seu_sample_subset[['batch']]) == 'B5'){
    multiplet_rate <- 0.023
  }
  if(unique(seu_sample_subset[['batch']]) == 'B6'){
    multiplet_rate <- 0.076
  }
  if(unique(seu_sample_subset[['batch']]) == 'B7'){
    multiplet_rate <- 0.031
  }
  if(unique(seu_sample_subset[['batch']]) == 'B8'){
    multiplet_rate <- 0.039
  }
  if(unique(seu_sample_subset[['batch']]) == 'B9'){
    multiplet_rate <- 0.092
  }
  if(unique(seu_sample_subset[['batch']]) == 'B10'){
    multiplet_rate <- 0.076
  }
  
  # Pre-process seurat object with standard seurat workflow --- 
  sample <- NormalizeData(seu_sample_subset, verbose = FALSE)
  sample <- FindVariableFeatures(sample, verbose = FALSE)
  sample <- ScaleData(sample, verbose = FALSE)
  #sample <- SCTransform(seu_sample_subset, verbose = FALSE)
  #sample <- RunPCA(sample, nfeatures.print = 10, verbose = FALSE)
  sample <- RunPCA(sample, verbose = FALSE)
  
  # Find significant PCs
  stdv <- sample[["pca"]]@stdev
  percent_stdv <- (stdv/sum(stdv)) * 100
  cumulative <- cumsum(percent_stdv)
  co1 <- which(cumulative > 90 & percent_stdv < 5)[1] 
  co2 <- sort(which((percent_stdv[1:length(percent_stdv) - 1] - 
                       percent_stdv[2:length(percent_stdv)]) > 0.1), 
              decreasing = T)[1] + 1
  min_pc <- min(co1, co2)
  
  # Finish pre-processing with min_pc
  sample <- RunUMAP(sample, dims = 1:min_pc, verbose = FALSE)
  sample <- FindNeighbors(object = sample, dims = 1:min_pc, verbose = FALSE)              
  sample <- FindClusters(object = sample, resolution = 0.1, verbose = FALSE)
  
  # pK identification (no ground-truth) 
  #introduces artificial doublets in varying props, merges with real data set and 
  # preprocesses the data + calculates the prop of artficial neighrest neighbours, 
  # provides a list of the proportion of artificial nearest neighbours for varying
  # combinations of the pN and pK
  sweep_list <- paramSweep(sample, PCs = 1:min_pc, sct = FALSE)   
  sweep_stats <- summarizeSweep(sweep_list)
  bcmvn <- find.pK(sweep_stats) # computes a metric to find the optimal pK value (max mean variance normalised by modality coefficient)
  # Optimal pK is the max of the bimodality coefficient (BCmvn) distribution
  optimal.pk <- bcmvn %>% 
    dplyr::filter(BCmetric == max(BCmetric)) %>%
    dplyr::select(pK)
  optimal.pk <- as.numeric(as.character(optimal.pk[[1]]))
  
  ## Homotypic doublet proportion estimate
  annotations <- sample@meta.data$seurat_clusters # use the clusters as the user-defined cell types
  homotypic.prop <- modelHomotypic(annotations) # get proportions of homotypic doublets
  
  #multiply by number of cells to get the number of expected multiplets
  nExp.poi <- round(multiplet_rate * nrow(sample@meta.data)) 
  nExp.poi.adj <- round(nExp.poi * (1 - homotypic.prop)) # expected number of doublets
  
  # run DoubletFinder
  sample <- doubletFinder(seu = sample, 
                          PCs = 1:min_pc, 
                          pK = optimal.pk, #the neighborhood size used to compute the number of artificial nearest neighbours
                          nExp = nExp.poi.adj) #number of expected real doublets
  
  # change name of metadata column with Singlet/Doublet information
  colnames(sample@meta.data)[grepl('DF.classifications.*', colnames(sample@meta.data))] <- "doublet_finder"
  
  #DimPlot(sample, group.by = "doublet_finder")
  
  #get the metadata column with singlet, doublet info
  double_finder_res <- sample@meta.data['doublet_finder'] 
  #add the cell IDs as new column to be able to merge correctly
  double_finder_res <- rownames_to_column(double_finder_res, "row_names") 
  return(double_finder_res)
}

data.split <- SplitObject(filteredData, split.by = "batch")
data.split <- lapply(data.split, run_doubletfinder_custom)

sglt_dblt_metadata <- data.frame(bind_rows(data.split)) # merge to a single dataframe
rownames(sglt_dblt_metadata) <- sglt_dblt_metadata$row_names # assign cell IDs to row names to ensure match
sglt_dblt_metadata$row_names <- NULL
head(sglt_dblt_metadata)
filteredData <- AddMetaData(filteredData, sglt_dblt_metadata, col.name = 'doublet_finder')

rm(data.split, feats, sglt_dblt_metadata)

# Check how doublets singlets differ in QC measures per sample.
VlnPlot(filteredData, group.by = 'batch', split.by = "doublet_finder",
        features = c("nFeature_RNA", "nCount_RNA", "percent_ribo"), 
        ncol = 3, pt.size = 0) + theme(legend.position = 'right')

```
Plot showing higher counts and featured genes in doublets than singlets. 


```{r Removing doublets and genes not present in both datasets}
filteredData <- filteredData[, filteredData@meta.data[, "doublet_finder"] == "Singlet"]

onlyLM <- subset(filteredData, type == "LM")
onlyPT <- subset(filteredData, orig.ident %in% c("3_JEPDX1_7A", "3_JEPDX1_7B", "4_JEPDX2_5", "7_CHPDX2_2", "8_JEPDX3_5"))
selectFeat <- rownames(onlyLM)[Matrix::rowSums(onlyLM[["RNA"]]$counts) > 1]
filteredData <- subset(filteredData, features = selectFeat)
selectFeat <- rownames(onlyPT)[Matrix::rowSums(onlyPT[["RNA"]]$counts) > 1]
filteredData <- subset(filteredData, features = selectFeat)

rm(onlyLM, onlyPT, selectFeat)
saveRDS(filteredData, file = "SeuratObject_filteredData.rds")
filteredData <- readRDS("SeuratObject_filteredData.rds")

dim(filteredData)
table(filteredData$type_mouse)
table(filteredData$PDX_type)
table(filteredData$type)
```
Dataset size after filtering
Filtering done:
*Remove doublets. (amount of doublets filtered is based of predicted doublets, which increase with batch size. Liver metastasis sampels wer done with sample 1-4 in one batch and 5-8 in one. All primary tumor samples were run in individual, larger batches, leading to a higher frequency of doublets as well.)

##########################################################################################################
##Preparation of data, Integration
Normalizing data: Each cell has a different total numbers of reads. Normalizing compensates for that.  
Finding variable features: Finding genes with high variance.
Scaling data: Centers and equalizes the value of each genes. Sets mean = 0 and variance = 1.
PCA: Weights the found variable features to capture as much variation as possible.

Not using SCTransform: https://www.reddit.com/r/bioinformatics/comments/1kdag9p/seurat_v5_sctransform_deg_analyses_and/
```{r Integrating and processing}
processedData <- filteredData
processedData[["RNA"]] <- split(processedData[["RNA"]], f = filteredData$batch)

processedData <- NormalizeData(processedData, verbose = FALSE)
processedData <- FindVariableFeatures(processedData, verbose = FALSE)
processedData <- ScaleData(processedData, verbose = FALSE)
processedData <- RunPCA(processedData, verbose = FALSE)

#Integrating
integratedData <- IntegrateLayers(object = processedData, method = HarmonyIntegration, group.by.vars = "batch", assay.use = "RNA", orig.reduction = "pca", new.reduction = "harmony", verbose = FALSE)
integratedData[["RNA"]] <- JoinLayers(integratedData[["RNA"]])

#objectList <- SplitObject(filteredData, split.by = "batch")
#objectList <- lapply(objectList, function(serObj){
#  serObjSCT <- SCTransform(serObj, return.only.var.genes = FALSE, verbose = FALSE)
#  #serObjSCT <- RunPCA(serObjSCT, assay = "SCT", verbose = FALSE)
#  serObjSCT
#})

#varFeat1 <- VariableFeatures(objectList[[1]])
#varFeat2 <- VariableFeatures(objectList[[2]])
#varFeat3 <- VariableFeatures(objectList[[3]])
#varFeat4 <- VariableFeatures(objectList[[4]])
#varFeat5 <- VariableFeatures(objectList[[5]])
#varFeat6 <- VariableFeatures(objectList[[6]])
#varFeat7 <- VariableFeatures(objectList[[7]])
#varFeat8 <- VariableFeatures(objectList[[8]])
#varFeat9 <- VariableFeatures(objectList[[9]])
#varFeat10 <- VariableFeatures(objectList[[10]])
#varFeat <- c(varFeat1, varFeat2, varFeat3, varFeat4, varFeat5, varFeat6, varFeat7, varFeat8, varFeat9, varFeat10)
#varFeat <- unique(varFeat)

#features <- SelectIntegrationFeatures(object.list = objectList, nfeatures = 3000)
#processedData <- merge(objectList[[1]], y = objectList[-1])
#VariableFeatures(processedData) <- features

#processedData <- RunPCA(processedData, assay = "SCT", verbose = FALSE)
#integratedData <- harmony::RunHarmony(object = processedData, group.by.vars = "batch", reduction = "pca", #assay.use = "SCT", reduction.save = "harmony", project.dim = FALSE)
#integratedData[["RNA"]] <- JoinLayers(integratedData[["RNA"]])



saveRDS(processedData, file = "SeuratObject_processedData.rds")
saveRDS(integratedData, file = "SeuratObject_integratedData.rds")

integratedData <- readRDS("SeuratObject_integratedData.rds")

#Printing components of PCA
wrap_plots(
  DimPlot(processedData, reduction = "pca", group.by = "PDX_type", dims = 1:2),
  DimPlot(processedData, reduction = "pca", group.by = "PDX_type", dims = 3:4),
  DimPlot(processedData, reduction = "pca", group.by = "PDX_type", dims = 5:6),
  ncol = 3
) + plot_layout(guides = "collect")

```
Showing the variance captured by first 6 PCA dimensions. (Maybe not relevant to show?) 
```{r PCA components}
#What genes contribute to PCs
VizDimLoadings(processedData, dims = 1:4, reduction = "pca", ncol = 2, balanced = T)
```
Showing what genes contributed first 5 PCA dimensions. (Relevant?)

##Visualising using UMAP and tSNE
UMAP is better for visualizing variability between cell clusters. tSNE is better for visualizing in-group variability. 
```{r UMAP of Integrated/non}
#UMAP on integrated data
data.umap <- RunUMAP(integratedData, dims = 1:15, reduction = "harmony", reduction.name = "umap_harmony", verbose = FALSE, assay = "RNA")
##UMAP on NonIntegrated data
#nonIntData <- RunUMAP(processedData, dims = 1:15, reduction = "pca", verbose = FALSE)

##tSNE on integrated data
#data.tsne <- RunTSNE(
#  integratedData,
#  reduction = "harmony", dims = 1:15,
#  perplexity = 30,
#  max_iter = 1000,
#  theta = 0.5,
#  eta = 200,
#  num_threads = 0,
#  verbose = FALSE
#)
##tSNE on NonIntegrated data
#data.tsneNonInt <- RunTSNE(
#  processedData,
#  reduction = "pca", dims = 1:15,
#  perplexity = 30,
#  max_iter = 1000,
#  theta = 0.5,
#  eta = 200,
#  num_threads = 0,
#  vebrose = FALSE
#)

cols_type <- c("PT" = "#6BAED6", "LM" = "#E6550D")
cols_PDX_type <- c("LM LU-NB-1" = "#084594", "LM LU-NB-2" = "#006D2C", "LM LU-NB-3" = "#A63603", "PT LU-NB-1" = "#2d9ffc", "PT LU-NB-2" = "#74C476", "PT LU-NB-3" = "#FD8D3C")
cols_PDX <- c("LU-NB-1" = "#2d9ffc", "LU-NB-2" = "#74C476", "LU-NB-3" = "#FD8D3C")
cols_PDX_LM <- c("LU-NB-1" = "#084594", "LU-NB-2" = "#006D2C", "LU-NB-3" = "#A63603")
#clustCol <- DiscretePalette_scCustomize(num_colors = 11, palette = "glasbey")
clustCol <- brewer.pal(n = 12, name = "Paired")
cols_clust <- c("1" = clustCol[1], "2" = clustCol[2], "3" = clustCol[3], "4" = clustCol[4], "5" = clustCol[5], "6" = clustCol[6], "7" = clustCol[7], "8" = clustCol[8], "9" = clustCol[9], "10" = clustCol[10])


##Summorize based on mouse
#wrap_plots(
#  DimPlot(nonIntData, reduction = "umap", group.by = "PDX_type", cols = cols_PDX_type) + ggtitle("NonIntegrated #UMAP"),
#  DimPlot(data.umap, reduction = "umap_harmony", group.by = "PDX_type", cols = cols_PDX_type) + #ggtitle("Harmony UMAP"),
#  DimPlot(nonIntData, reduction = "umap", group.by = "type", cols = cols_type) + ggtitle("NonIntegrated UMAP"),
#  DimPlot(data.umap, reduction = "umap_harmony", group.by = "type", cols = cols_type) + ggtitle("Harmony #UMAP"),  
#  ncol = 2
#) + plot_layout(guides = "collect")
```
Showing UMAP based on PDX and type before and after compensating for batch effect(integrating)
```{r tSNE of Integrated/Non}
#wrap_plots(
#  DimPlot(data.tsneNonInt, reduction = "tsne", group.by = "PDX_type", cols = cols_PDX_type) + #ggtitle("NonIntegrated tSNE"),
#  DimPlot(data.tsne, reduction = "tsne", group.by = "PDX_type", cols = cols_PDX_type) + ggtitle("Harmony #tSNE"),
#  DimPlot(data.tsneNonInt, reduction = "tsne", group.by = "type", cols = cols_type) + ggtitle("NonIntegrated #tSNE"),
#  DimPlot(data.tsne, reduction = "tsne", group.by = "type", cols = cols_type) + ggtitle("Harmony tSNE"),
#  ncol = 2
#) + plot_layout(guides = "collect")
#rm(data.tsneNonInt, data.tsne, nonIntData, integratedData)
```
Showing tSNE based on PDX and type and before and after compensating for batch effect(integrating)
```{r UMAP of individual lines and LM}
#onlyLM.umap <- subset(data.umap, type == "LM")
#onlyPDX1.umap <- subset(data.umap, PDX == "LU-NB-1" & type == "PT")
#onlyPDX2.umap <- subset(data.umap, PDX == "LU-NB-2" & type == "PT")
#onlyPDX3.umap <- subset(data.umap, PDX == "LU-NB-3" & type == "PT")
#wrap_plots(
#  DimPlot(onlyPDX1.umap, reduction = "umap_harmony", group.by = "type", cols = "#2d9ffc") + NoLegend() + #ggtitle("PT LU-NB-1"),
#  DimPlot(onlyPDX2.umap, reduction = "umap_harmony", group.by = "type", cols = "#74C476") + NoLegend() + #ggtitle("PT LU-NB-2"),
#  DimPlot(onlyPDX3.umap, reduction = "umap_harmony", group.by = "type", cols = "#FD8D3C") + NoLegend() + #ggtitle("PT LU-NB-3"),
#  DimPlot(onlyLM.umap, reduction = "umap_harmony", group.by = "PDX", cols = cols_PDX_LM) + ggtitle("LM"),
#  ncol = 2
#) + plot_layout(guides = "collect")
#
#rm(onlyLM.umap, onlyPDX1.umap, onlyPDX2.umap, onlyPDX3.umap)
```


##Clustering
```{r Clustering }
#Clustering
data.umap <- FindNeighbors(data.umap, dims = 1:15, verbose = FALSE, reduction = "harmony")
data.umap <- FindClusters(data.umap, resolution = 0.30, verbose = FALSE)

#data.umap$seurat_clusters <- as.character(as.integer(data.umap$seurat_clusters) + 1)
#data.umap$seurat_clusters <- factor(data.umap$seurat_clusters, levels = 1:11)

#data.tsne <- FindNeighbors(data.tsne, dims = 1:20, verbose = FALSE)
#data.tsne <- FindClusters(data.tsne, resolution = 0.3, verbose = FALSE)

#Visualising if calculated clusters correspond to celltypes

#Setting cluster names 1-12
new.cluster.ids <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10") 
names(new.cluster.ids) <- levels(data.umap)
data.umap <- RenameIdents(data.umap, new.cluster.ids)
data.umap$seurat_clusters <- Idents(data.umap)

Fig1a <- DimPlot(data.umap, reduction = "umap_harmony", group.by = "type", cols = cols_type, order = c("LM", "PT")) + ggtitle("Per Location") + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1),legend.position = "inside", legend.position.inside = c(0.8, 0.8)) + ylab("UMAP_2") + xlab("UMAP_1")
#DimPlot(data.umap, reduction = "umap_harmony", group.by = "PDX", cols = cols_PDX) + ggtitle("Cell line"), 
Fig1b <- DimPlot(data.umap, reduction = "umap_harmony", group.by = "seurat_clusters", cols = cols_clust) + ggtitle("Per Cluster") + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1)) + ylab("UMAP_2") + xlab("UMAP_1")

#Create a Dimplot grid of 2x3 with PDX as columns and rows as type, plotting only cells agreeing with boths. With PDX as column label and type as row label. Styled as the DimPlots above.
#Create the same dimplot grid but styled with black borders and axis titles and no axis ticks or text. With column titles and row titles, but not individual plot titles. 
wrap_plots(
  DimPlot(subset(data.umap, type == "PT" & PDX == "LU-NB-1"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend() + ggtitle("LU-NB-1"),
  DimPlot(subset(data.umap, type == "PT" & PDX == "LU-NB-2"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend() + ggtitle("LU-NB-2"),
  DimPlot(subset(data.umap, type == "PT" & PDX == "LU-NB-3"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend() + ggtitle("LU-NB-3"),
  DimPlot(subset(data.umap, type == "LM" & PDX == "LU-NB-1"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend(),
  DimPlot(subset(data.umap, type == "LM" & PDX == "LU-NB-2"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend(),
  DimPlot(subset(data.umap, type == "LM" & PDX == "LU-NB-3"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend(),
  ncol = 3
) + plot_layout(guides = "collect") & theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1)) & theme(legend.position = "none")

# Q: Can one make a grid of dimplots that share borders with eachother? With column titles and row titles, but not individual plot titles?
# A: Yes, using wrap_plots and plot_layout from the patchwork package. You can create a grid of dimplots and customize the theme to have shared borders and specific titles for columns and rows. In the example above, I created a 2x3 grid of dimplots based on PDX and type, with customized themes to achieve the desired appearance.
# Q: The one you created above did not have shared borders, how can one create a this grid in another way?
# A: To create a grid of dimplots with shared borders, you can use the cowplot package in R. Here's an example of how to do this:
cowplot::plot_grid(
  DimPlot(subset(data.umap, type == "PT" & PDX == "LU-NB-1"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend() + ggtitle("LU-NB-1") + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1), plot.margin = unit(c(0, 0, 0, 0), "cm")) + ylab("PT") + xlab(NULL),
  DimPlot(subset(data.umap, type == "PT" & PDX == "LU-NB-2"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend() + ggtitle("LU-NB-2") + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1), plot.margin = unit(c(0, 0, 0, 0), "cm")) + ylab(NULL) + xlab(NULL),
  DimPlot(subset(data.umap, type == "PT" & PDX == "LU-NB-3"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend() + ggtitle("LU-NB-3") + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1), plot.margin = unit(c(0, 0, 0, 0), "cm")) + ylab(NULL) + xlab(NULL),
  DimPlot(subset(data.umap, type == "LM" & PDX == "LU-NB-1"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend() + ggtitle(NULL) + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1), plot.margin = unit(c(0, 0, 0, 0), "cm")) + ylab("LM") + xlab(NULL),
  DimPlot(subset(data.umap, type == "LM" & PDX == "LU-NB-2"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend() + ggtitle(NULL) + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1), plot.margin = unit(c(0, 0, 0, 0), "cm")) + ylab(NULL) + xlab(NULL),
  DimPlot(subset(data.umap, type == "LM" & PDX == "LU-NB-3"), reduction = "umap_harmony", group.by = "type", cols = cols_type) + NoLegend() + ggtitle(NULL) + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1), plot.margin = unit(c(0, 0, 0, 0), "cm")) + ylab(NULL) + xlab(NULL),
  ncol = 3,
  align = 'hv',
  axis = 'tblr'
)

data.umap$PDX_type <- factor(data.umap$PDX_type, levels = c("PT LU-NB-1", "PT LU-NB-2", "PT LU-NB-3", "LM LU-NB-1", "LM LU-NB-2", "LM LU-NB-3"))

DimPlot(data.umap, reduction = "umap_harmony", group.by = "type", split.by = "PDX_type", cols = cols_type, ncol = 3) + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1), legend.position = "none") + ylab("UMAP_2") + xlab("UMAP_1")

saveRDS(Fig1a, file = "Fig1a.rds")
saveRDS(Fig1b, file = "Fig1b.rds")
```
Visualizing of calculated clusters and how they correspond to celltype
```{r seperated UMAPs}
#DimPlot(data.umap, reduction = "umap_harmony", group.by = "type", cols = cols_type, order = c("LM", "PT")) + #ggtitle("Location")
#  #DimPlot(data.umap, reduction = "umap_harmony", group.by = "PDX", cols = cols_PDX) + ggtitle("Cell line"), 
#DimPlot(data.umap, reduction = "umap_harmony", group.by = "seurat_clusters", cols = cols_clust) + #ggtitle("Clustered")
```


```{r Barplot of clusters in PDXs}
onlyPTPDX1 <- subset(data.umap, PDX == "LU-NB-1" & type == "PT")
onlyPTPDX2 <- subset(data.umap, PDX == "LU-NB-2" & type == "PT")
onlyPTPDX3 <- subset(data.umap, PDX == "LU-NB-3" & type == "PT")
onlyLMPDX1 <- subset(data.umap, PDX == "LU-NB-1" & type == "LM")
onlyLMPDX2 <- subset(data.umap, PDX == "LU-NB-2" & type == "LM")
onlyLMPDX3 <- subset(data.umap, PDX == "LU-NB-3" & type == "LM")

CelllineType <- c(rep("PT LU-NB-1", 10), rep("LM LU-NB-1", 10), rep("PT LU-NB-2", 10), rep("LM LU-NB-2", 10), rep("PT LU-NB-3", 10), rep("LM LU-NB-3", 10))
Clusters <- c(rep(1:10, 6))
Percentage <- c(c(100*length(WhichCells(onlyPTPDX1, expression = seurat_clusters == "1"))/length(colnames(onlyPTPDX1)), 100*length(WhichCells(onlyPTPDX1, expression = seurat_clusters == "2"))/length(colnames(onlyPTPDX1)), 100*length(WhichCells(onlyPTPDX1, expression = seurat_clusters == "3"))/length(colnames(onlyPTPDX1)), 100*length(WhichCells(onlyPTPDX1, expression = seurat_clusters == "4"))/length(colnames(onlyPTPDX1)), 100*length(WhichCells(onlyPTPDX1, expression = seurat_clusters == "5"))/length(colnames(onlyPTPDX1)), 100*length(WhichCells(onlyPTPDX1, expression = seurat_clusters == "6"))/length(colnames(onlyPTPDX1)), 100*length(WhichCells(onlyPTPDX1, expression = seurat_clusters == "7"))/length(colnames(onlyPTPDX1)), 100*length(WhichCells(onlyPTPDX1, expression = seurat_clusters == "8"))/length(colnames(onlyPTPDX1)), 100*length(WhichCells(onlyPTPDX1, expression = seurat_clusters == "9"))/length(colnames(onlyPTPDX1)), 100*length(WhichCells(onlyPTPDX1, expression = seurat_clusters == "10"))/length(colnames(onlyPTPDX1))),
              c(100*length(WhichCells(onlyLMPDX1, expression = seurat_clusters == "1"))/length(colnames(onlyLMPDX1)), 100*length(WhichCells(onlyLMPDX1, expression = seurat_clusters == "2"))/length(colnames(onlyLMPDX1)), 100*length(WhichCells(onlyLMPDX1, expression = seurat_clusters == "3"))/length(colnames(onlyLMPDX1)), 100*length(WhichCells(onlyLMPDX1, expression = seurat_clusters == "4"))/length(colnames(onlyLMPDX1)), 100*length(WhichCells(onlyLMPDX1, expression = seurat_clusters == "5"))/length(colnames(onlyLMPDX1)), 100*length(WhichCells(onlyLMPDX1, expression = seurat_clusters == "6"))/length(colnames(onlyLMPDX1)), 100*length(WhichCells(onlyPTPDX1, expression = seurat_clusters == "7"))/length(colnames(onlyLMPDX1)), 100*length(WhichCells(onlyLMPDX1, expression = seurat_clusters == "8"))/length(colnames(onlyLMPDX1)), 100*length(WhichCells(onlyLMPDX1, expression = seurat_clusters == "9"))/length(colnames(onlyLMPDX1)), 100*length(WhichCells(onlyLMPDX1, expression = seurat_clusters == "10"))/length(colnames(onlyLMPDX1))),
              c(100*length(WhichCells(onlyPTPDX2, expression = seurat_clusters == "1"))/length(colnames(onlyPTPDX2)), 100*length(WhichCells(onlyPTPDX2, expression = seurat_clusters == "2"))/length(colnames(onlyPTPDX2)), 100*length(WhichCells(onlyPTPDX2, expression = seurat_clusters == "3"))/length(colnames(onlyPTPDX2)), 100*length(WhichCells(onlyPTPDX2, expression = seurat_clusters == "4"))/length(colnames(onlyPTPDX2)), 100*length(WhichCells(onlyPTPDX2, expression = seurat_clusters == "5"))/length(colnames(onlyPTPDX2)), 100*length(WhichCells(onlyPTPDX2, expression = seurat_clusters == "6"))/length(colnames(onlyPTPDX2)), 100*length(WhichCells(onlyPTPDX2, expression = seurat_clusters == "7"))/length(colnames(onlyPTPDX2)), 100*length(WhichCells(onlyPTPDX2, expression = seurat_clusters == "8"))/length(colnames(onlyPTPDX2)), 100*length(WhichCells(onlyPTPDX2, expression = seurat_clusters == "9"))/length(colnames(onlyPTPDX2)), 100*length(WhichCells(onlyPTPDX2, expression = seurat_clusters == "10"))/length(colnames(onlyPTPDX2))), 
              c(100*length(WhichCells(onlyLMPDX2, expression = seurat_clusters == "1"))/length(colnames(onlyLMPDX2)), 100*length(WhichCells(onlyLMPDX2, expression = seurat_clusters == "2"))/length(colnames(onlyLMPDX2)), 100*length(WhichCells(onlyLMPDX2, expression = seurat_clusters == "3"))/length(colnames(onlyLMPDX2)), 100*length(WhichCells(onlyLMPDX2, expression = seurat_clusters == "4"))/length(colnames(onlyLMPDX2)), 100*length(WhichCells(onlyLMPDX2, expression = seurat_clusters == "5"))/length(colnames(onlyLMPDX2)), 100*length(WhichCells(onlyLMPDX2, expression = seurat_clusters == "6"))/length(colnames(onlyLMPDX2)), 100*length(WhichCells(onlyLMPDX2, expression = seurat_clusters == "7"))/length(colnames(onlyLMPDX2)), 100*length(WhichCells(onlyLMPDX2, expression = seurat_clusters == "8"))/length(colnames(onlyLMPDX2)), 100*length(WhichCells(onlyLMPDX2, expression = seurat_clusters == "9"))/length(colnames(onlyLMPDX2)), 100*length(WhichCells(onlyLMPDX2, expression = seurat_clusters == "10"))/length(colnames(onlyLMPDX2))), 
              c(100*length(WhichCells(onlyPTPDX3, expression = seurat_clusters == "1"))/length(colnames(onlyPTPDX3)), 100*length(WhichCells(onlyPTPDX3, expression = seurat_clusters == "2"))/length(colnames(onlyPTPDX3)), 100*length(WhichCells(onlyPTPDX3, expression = seurat_clusters == "3"))/length(colnames(onlyPTPDX3)), 100*length(WhichCells(onlyPTPDX3, expression = seurat_clusters == "4"))/length(colnames(onlyPTPDX3)), 100*length(WhichCells(onlyPTPDX3, expression = seurat_clusters == "5"))/length(colnames(onlyPTPDX3)), 100*length(WhichCells(onlyPTPDX3, expression = seurat_clusters == "6"))/length(colnames(onlyPTPDX3)), 100*length(WhichCells(onlyPTPDX3, expression = seurat_clusters == "7"))/length(colnames(onlyPTPDX3)), 100*length(WhichCells(onlyPTPDX3, expression = seurat_clusters == "8"))/length(colnames(onlyPTPDX3)), 100*length(WhichCells(onlyPTPDX3, expression = seurat_clusters == "9"))/length(colnames(onlyPTPDX3)), 100*length(WhichCells(onlyPTPDX3, expression = seurat_clusters == "10"))/length(colnames(onlyPTPDX3))),
              c(100*length(WhichCells(onlyLMPDX3, expression = seurat_clusters == "1"))/length(colnames(onlyLMPDX3)), 100*length(WhichCells(onlyLMPDX3, expression = seurat_clusters == "2"))/length(colnames(onlyLMPDX3)), 100*length(WhichCells(onlyLMPDX3, expression = seurat_clusters == "3"))/length(colnames(onlyLMPDX3)), 100*length(WhichCells(onlyLMPDX3, expression = seurat_clusters == "4"))/length(colnames(onlyLMPDX3)), 100*length(WhichCells(onlyLMPDX3, expression = seurat_clusters == "5"))/length(colnames(onlyLMPDX3)), 100*length(WhichCells(onlyLMPDX3, expression = seurat_clusters == "6"))/length(colnames(onlyLMPDX3)), 100*length(WhichCells(onlyLMPDX3, expression = seurat_clusters == "7"))/length(colnames(onlyLMPDX3)), 100*length(WhichCells(onlyLMPDX3, expression = seurat_clusters == "8"))/length(colnames(onlyLMPDX3)), 100*length(WhichCells(onlyLMPDX3, expression = seurat_clusters == "9"))/length(colnames(onlyLMPDX3)), 100*length(WhichCells(onlyLMPDX3, expression = seurat_clusters == "10"))/length(colnames(onlyLMPDX3))))

barplotData <- data.frame(CelllineType, Clusters, Percentage)
barplotData$Clusters <- factor(barplotData$Clusters, levels = 1:10)
barplotData$CelllineType <- factor(barplotData$CelllineType, levels = c("PT LU-NB-1", "LM LU-NB-1", "PT LU-NB-2", "LM LU-NB-2", "PT LU-NB-3", "LM LU-NB-3"))

Fig1c <- ggplot(barplotData, aes(fill=Clusters, y=Percentage, x=CelllineType, size=8)) + 
  geom_bar(position = "fill", stat = "identity", width = 0.9) + 
  ggtitle("Percentage of Clusters per PDX and Location") +
  scale_fill_manual(values = cols_clust) +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0, 0)) +
  scale_x_discrete(limits = c("PT LU-NB-1", "LM LU-NB-1", "", "PT LU-NB-2", "LM LU-NB-2", "", "PT LU-NB-3", "LM LU-NB-3"), labels = c("PT LU-NB-1", "LM LU-NB-1", "", "PT LU-NB-2", "LM LU-NB-2", "", "PT LU-NB-3", "LM LU-NB-3")) +
  #annotate(geom = "text", x = 1, y = -0.02, label = "PT", size = 3) +
  #annotate(geom = "text", x = 2, y = -0.02, label = "LM", size = 3) +
  #annotate(geom = "text", x = 4, y = -0.02, label = "PT", size = 3) +
  #annotate(geom = "text", x = 5, y = -0.02, label = "LM", size = 3) +
  #annotate(geom = "text", x = 7, y = -0.02, label = "PT", size = 3) +
  #annotate(geom = "text", x = 8, y = -0.02, label = "LM", size = 3) +
  #theme_minimal() +
  geom_text(aes(label = ifelse(Percentage > 3, round(Percentage, 2), "")), 
            position = position_fill(vjust = 0.5), color =  "black", size = 4) +
  theme_minimal() +
  theme(
  legend.position = "right",
  legend.title = element_blank(),
  axis.title.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.y = element_blank(),
  axis.line.y = element_line(color = "black", linewidth = 1.5, lineend = "square"),
  axis.ticks.y = element_line(color = "black", linewidth = 1.5),
  axis.ticks.length.y = unit(0.2, "cm"),
  axis.line.x = element_blank(),
  axis.text.x = element_text(angle = 45, hjust = 1, size = 15),
  axis.text.y = element_text(size = 15),
  panel.grid = element_blank(),
  text = element_text(family = "sans", size = 12, color = "black")
)

saveRDS(Fig1c, file = "Fig1c.rds")

rm(onlyPTPDX1, onlyPTPDX2, onlyPTPDX3, onlyLMPDX1, onlyLMPDX2, onlyLMPDX3, CelllineType, Fraction, Clusters, barplotData)
```


##########################################################################################################
##Investigating clusters
```{r Finding markers/cluster + heatmap}
#Finding cluster markers and plotting in heatmap
#data.umap <- PrepSCTFindMarkers(data.umap, assay = "SCT", verbose = TRUE)
clusterMarkers <- FindAllMarkers(data.umap, group.by = "seurat_clusters")
clusterMarkers %>% 
  group_by(cluster) %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  slice_head(n = 10) %>%
  #slice_max(avg_log2FC, n = 10) %>%
  #ungroup() %>%
  #group_by(gene) %>%
  #slice_max(avg_log2FC, n = 1) %>%
  ungroup() -> top10Cluster
#clusterHeatmap <- ScaleData(data.umap, features = top10Cluster$gene)
#clusterHeatmap <- GetResidual(data.umap, features = top10Cluster$gene)
#DoHeatmap(data.umap, features = top10Cluster$gene)
```
Heatmap av top 10 gener i varje cluster

```{r ComplexHeatmap}
top10AdrnGenes <- c("RBP1", "NPW", "PCSK1N", "GAL", "NBEA", "LINGO2", "TMSB10", "STMN2", "S100A1", "CDKN3", "RRM2", "UBE2C", "BIRC5", "PBK", "NUSAP1", "EPB41L3", "DOCK3")
top10MesGenes <- c("VIM", "S100A6", "APOE", "LGALS1", "MT2A", "S100A4", "HLA-B", "IGF2BP2", "ALX1")
top10ProlifGenes <- c("CCNB1", "CCNB2", "PTTG1", "BIRC5", "CDKN3", "DLGAP5", "UBE2C", "CENPA", "NUF2", "TPX2", "RRM2", "CDK1", "GINS2", "TYMS", "UBE2T", "PCLAF", "CENPW", "PBK", "NUSAP1")

top10Cluster <- unique(top10Cluster$gene)
clusterHeatmap <- ScaleData(data.umap, features = top10Cluster)
heatmapSer0 <- subset(clusterHeatmap, features = top10Cluster, seurat_clusters == "1")
heatmapMatr0 <- GetAssayData(heatmapSer0, assay = "RNA", layer = "scale.data")
heatmapMatr0 <- heatmapMatr0[order(factor(rownames(heatmapMatr0), levels = top10Cluster)), ]
heatmapAnn0 <- HeatmapAnnotation(Cluster = heatmapSer0$seurat_clusters, PDX = heatmapSer0$PDX, Location = heatmapSer0$type, col = list(Cluster = cols_clust, PDX = cols_PDX, Location = cols_type), annotation_label = c(" ", " ", " "), show_legend = FALSE)
heatmapAnnR <- rowAnnotation(Geneinf = ifelse(rownames(heatmapMatr0) %in% top10AdrnGenes, "ADRN", ifelse(rownames(heatmapMatr0) %in% top10MesGenes, "MES", ifelse(rownames(heatmapMatr0) %in% top10ProlifGenes, "Prolif", "."))), show_legend = FALSE, col = list(Geneinf = c("ADRN" = "orange", "MES" = "navy", "Prolif" = "green", "." = "white")), annotation_label = c(" "), simple_anno_size = unit(2, "mm"))
col_fun = colorRamp2(c(-2, 0, 2), c("#FF00FF", "#000000", "#FFFF00"))
h0 <- Heatmap(heatmapMatr0, name = "clusterHeatmap", cluster_rows = FALSE, cluster_columns = FALSE, show_column_dend = FALSE, show_column_names = FALSE, show_row_dend = FALSE , top_annotation = heatmapAnn0, col = col_fun, use_raster = FALSE, column_title = "1", show_heatmap_legend = FALSE, row_names_gp = gpar(fontsize = 6))

heatmapSer0 <- subset(clusterHeatmap, features = top10Cluster, seurat_clusters == "2")
heatmapMatr0 <- GetAssayData(heatmapSer0, assay = "RNA", layer = "scale.data")
heatmapMatr0 <- heatmapMatr0[order(factor(rownames(heatmapMatr0), levels = top10Cluster)), ]
heatmapAnn0 <- HeatmapAnnotation(Cluster = heatmapSer0$seurat_clusters, PDX = heatmapSer0$PDX, Location = heatmapSer0$type, col = list(Cluster = cols_clust, PDX = cols_PDX, Location = cols_type), annotation_label = c(" ", " ", " "), show_legend = FALSE)
h1 <- Heatmap(heatmapMatr0, name = "clusterHeatmap", cluster_rows = FALSE, cluster_columns = FALSE, show_column_dend = FALSE, show_column_names = FALSE, show_row_dend = FALSE , top_annotation = heatmapAnn0, col = col_fun, use_raster = FALSE, column_title = "2", show_heatmap_legend = FALSE, row_names_gp = gpar(fontsize = 6))

heatmapSer0 <- subset(clusterHeatmap, features = top10Cluster, seurat_clusters == "3")
heatmapMatr0 <- GetAssayData(heatmapSer0, assay = "RNA", layer = "scale.data")
heatmapMatr0 <- heatmapMatr0[order(factor(rownames(heatmapMatr0), levels = top10Cluster)), ]
heatmapAnn0 <- HeatmapAnnotation(Cluster = heatmapSer0$seurat_clusters, PDX = heatmapSer0$PDX, Location = heatmapSer0$type, col = list(Cluster = cols_clust, PDX = cols_PDX, Location = cols_type), annotation_label = c(" ", " ", " "), show_legend = FALSE)
h2 <- Heatmap(heatmapMatr0, name = "clusterHeatmap", cluster_rows = FALSE, cluster_columns = FALSE, show_column_dend = FALSE, show_column_names = FALSE, show_row_dend = FALSE , top_annotation = heatmapAnn0, col = col_fun, use_raster = FALSE, column_title = "3", show_heatmap_legend = FALSE, row_names_gp = gpar(fontsize = 6))

heatmapSer0 <- subset(clusterHeatmap, features = top10Cluster, seurat_clusters == "4")
heatmapMatr0 <- GetAssayData(heatmapSer0, assay = "RNA", layer = "scale.data")
heatmapMatr0 <- heatmapMatr0[order(factor(rownames(heatmapMatr0), levels = top10Cluster)), ]
heatmapAnn0 <- HeatmapAnnotation(Cluster = heatmapSer0$seurat_clusters, PDX = heatmapSer0$PDX, type = heatmapSer0$type, col = list(Cluster = cols_clust, PDX = cols_PDX, type = cols_type), annotation_label = c(" ", " ", " "), show_legend = FALSE)
h3 <- Heatmap(heatmapMatr0, name = "clusterHeatmap", cluster_rows = FALSE, cluster_columns = FALSE, show_column_dend = FALSE, show_column_names = FALSE, show_row_dend = FALSE , top_annotation = heatmapAnn0, col = col_fun, use_raster = FALSE, column_title = "4", show_heatmap_legend = FALSE, row_names_gp = gpar(fontsize = 6))

heatmapSer0 <- subset(clusterHeatmap, features = top10Cluster, seurat_clusters == "5")
heatmapMatr0 <- GetAssayData(heatmapSer0, assay = "RNA", layer = "scale.data")
heatmapMatr0 <- heatmapMatr0[order(factor(rownames(heatmapMatr0), levels = top10Cluster)), ]
heatmapAnn0 <- HeatmapAnnotation(Cluster = heatmapSer0$seurat_clusters, PDX = heatmapSer0$PDX, Location = heatmapSer0$type, col = list(Cluster = cols_clust, PDX = cols_PDX, Location = cols_type), annotation_label = c(" ", " ", " "), show_legend = FALSE)
h4 <- Heatmap(heatmapMatr0, name = "clusterHeatmap", cluster_rows = FALSE, cluster_columns = FALSE, show_column_dend = FALSE, show_column_names = FALSE, show_row_dend = FALSE , top_annotation = heatmapAnn0, col = col_fun, use_raster = FALSE, column_title = "5", show_heatmap_legend = FALSE, row_names_gp = gpar(fontsize = 6))

heatmapSer0 <- subset(clusterHeatmap, features = top10Cluster, seurat_clusters == "6")
heatmapMatr0 <- GetAssayData(heatmapSer0, assay = "RNA", layer = "scale.data")
heatmapMatr0 <- heatmapMatr0[order(factor(rownames(heatmapMatr0), levels = top10Cluster)), ]
heatmapAnn0 <- HeatmapAnnotation(Cluster = heatmapSer0$seurat_clusters, PDX = heatmapSer0$PDX, Location = heatmapSer0$type, col = list(Cluster = cols_clust, PDX = cols_PDX, Location = cols_type), annotation_label = c(" ", " ", " "), show_legend = FALSE)
h5 <- Heatmap(heatmapMatr0, name = "clusterHeatmap", cluster_rows = FALSE, cluster_columns = FALSE, show_column_dend = FALSE, show_column_names = FALSE, show_row_dend = FALSE , top_annotation = heatmapAnn0, col = col_fun, use_raster = FALSE, column_title = "6", show_heatmap_legend = FALSE, row_names_gp = gpar(fontsize = 6))

heatmapSer0 <- subset(clusterHeatmap, features = top10Cluster, seurat_clusters == "7")
heatmapMatr0 <- GetAssayData(heatmapSer0, assay = "RNA", layer = "scale.data")
heatmapMatr0 <- heatmapMatr0[order(factor(rownames(heatmapMatr0), levels = top10Cluster)), ]
heatmapAnn0 <- HeatmapAnnotation(Cluster = heatmapSer0$seurat_clusters, PDX = heatmapSer0$PDX, Location = heatmapSer0$type, col = list(Cluster = cols_clust, PDX = cols_PDX, Location = cols_type), annotation_label = c(" ", " ", " "), show_legend = FALSE)
h6 <- Heatmap(heatmapMatr0, name = "clusterHeatmap", cluster_rows = FALSE, cluster_columns = FALSE, show_column_dend = FALSE, show_column_names = FALSE, show_row_dend = FALSE , top_annotation = heatmapAnn0, col = col_fun, use_raster = FALSE, column_title = "7", show_heatmap_legend = FALSE, row_names_gp = gpar(fontsize = 6))

heatmapSer0 <- subset(clusterHeatmap, features = top10Cluster, seurat_clusters == "8")
heatmapMatr0 <- GetAssayData(heatmapSer0, assay = "RNA", layer = "scale.data")
heatmapMatr0 <- heatmapMatr0[order(factor(rownames(heatmapMatr0), levels = top10Cluster)), ]
heatmapAnn0 <- HeatmapAnnotation(Cluster = heatmapSer0$seurat_clusters, PDX = heatmapSer0$PDX, Location = heatmapSer0$type, col = list(Cluster = cols_clust, PDX = cols_PDX, Location = cols_type), annotation_label = c(" ", " ", " "), show_legend = FALSE)
h7 <- Heatmap(heatmapMatr0, name = "clusterHeatmap", cluster_rows = FALSE, cluster_columns = FALSE, show_column_dend = FALSE, show_column_names = FALSE, show_row_dend = FALSE , top_annotation = heatmapAnn0, col = col_fun, use_raster = FALSE, column_title = "8", show_heatmap_legend = FALSE, row_names_gp = gpar(fontsize = 6))

heatmapSer0 <- subset(clusterHeatmap, features = top10Cluster, seurat_clusters == "9")
heatmapMatr0 <- GetAssayData(heatmapSer0, assay = "RNA", layer = "scale.data")
heatmapMatr0 <- heatmapMatr0[order(factor(rownames(heatmapMatr0), levels = top10Cluster)), ]
heatmapAnn0 <- HeatmapAnnotation(Cluster = heatmapSer0$seurat_clusters, PDX = heatmapSer0$PDX, Location = heatmapSer0$type, col = list(Cluster = cols_clust, PDX = cols_PDX, Location = cols_type), annotation_label = c(" ", " ", " "), show_legend = FALSE)
h8 <- Heatmap(heatmapMatr0, name = "clusterHeatmap", cluster_rows = FALSE, cluster_columns = FALSE, show_column_dend = FALSE, show_column_names = FALSE, show_row_dend = FALSE , top_annotation = heatmapAnn0, col = col_fun, use_raster = FALSE, column_title = "9", show_heatmap_legend = FALSE, row_names_gp = gpar(fontsize = 6))

heatmapSer0 <- subset(clusterHeatmap, features = top10Cluster, seurat_clusters == "10")
heatmapMatr0 <- GetAssayData(heatmapSer0, assay = "RNA", layer = "scale.data")
heatmapMatr0 <- heatmapMatr0[order(factor(rownames(heatmapMatr0), levels = top10Cluster)), ]
heatmapAnn0 <- HeatmapAnnotation(Cluster = heatmapSer0$seurat_clusters, PDX = heatmapSer0$PDX, Location = heatmapSer0$type, col = list(Cluster = cols_clust, PDX = cols_PDX, Location = cols_type), annotation_label = c("Cluster", "PDX", "Type"), show_legend = FALSE)
h9 <- Heatmap(heatmapMatr0, name = "clusterHeatmap", cluster_rows = FALSE, cluster_columns = FALSE, show_column_dend = FALSE, show_column_names = FALSE, show_row_dend = FALSE , top_annotation = heatmapAnn0, col = col_fun, use_raster = FALSE, column_title = "10", show_heatmap_legend = FALSE, row_names_gp = gpar(fontsize = 6), width = unit(3, "mm"), right_annotation = heatmapAnnR)


matrixLGND <- Legend(col_fun = col_fun, title = "Expression")
clusterLGND <- Legend(at = 1:10, title = "Clusters", legend_gp = gpar(fill = cols_clust))
cellLineLGND <- Legend(labels = c("LU-NB-1", "LU-NB-2", "LU-NB-3"), title = "PDX", legend_gp = gpar(fill = cols_PDX))
typeLGND <- Legend(labels = c("PT", "LM"), title = "Tumor type", legend_gp = gpar(fill = cols_type))
geneLGND <- Legend(labels = c("ADRN", "MES", "Prolif"), title = "Gene information", legend_gp = gpar(fill = c("ADRN" = "orange", "MES" = "navy", "Prolif" = "green")))
LGNDs <- packLegend(matrixLGND, clusterLGND, cellLineLGND, typeLGND, geneLGND)

fullHTMap <- h0 + h1 + h2 + h3 + h4 + h5 + h6 + h7 + h8 + h9
Fig1d <- draw(fullHTMap, ht_gap = unit(1, "mm"), annotation_legend_list = LGNDs)

saveRDS(Fig1d, file = "Fig1d.rds")

rm(h0, h1, h2, h3, h4, h5, h6, h7, h8, h9, heatmapSer0, heatmapMatr0, heatmapAnn0, matrixLGND, clusterLGND, cellLineLGND, typeLGND, LGNDs, fullHTMap, clusterHeatmap, geneLGND)
```


```{r Adr/mes genes}
#comparing clusters to reference genes for adrenergic, mesenchymal and proliferating cellstates
#adrn.genes <- c("PHOX2B","DBH","TH","DDC","CHGA","CHGB","SLC6A2","GATA3","HAND2","ISL1","ASCL1", "ZNF536", "PHOX2A", "HAND1", "ASCL1", "KLF13", "SOX11", "GATA2")
adrn.genes2 <- c("TH", "DBH", "CHGA", "PHOX2B", "GATA3", "MYCN", "PHOX2A")
#mes.genes <- c("VIM", "COL1A1","CXCL8", "TWIST2", "YAP1", "CD44", "COL3A1", "AXL", "FN1", "ACTA2", "S100A4", "RUNX2", "WWTR1", "IRF1", "PRRX1")
mes.genes2 <- c("CD44", "FN1", "COL1A1", "RUNX2", "WWTR1", "IRF1", "SNAI2", "VIM")
#prolif.genes <- c("MKI67","TOP2A","CENPF","UBE2C","HMGB2","TYMS", "PLK1", "E2F1", "FOXM1")


#FeaturePlot(data.umap, features = mes.genes2, ncol = 3)
#FeaturePlot(data.umap, features = adrn.genes2, ncol = 3)

```

```{r Calculate Adr/Mes score}
data.umap <- AddModuleScore(data.umap, features = list(adrn.genes2), name = "ADRN")
data.umap <- AddModuleScore(data.umap, features = list(mes.genes2), name = "MES")
#data.umap <- AddModuleScore(data.umap, features = list(prolif.genes), name = "Prolif")
#Plotting

p <- FeaturePlot(data.umap, features = c("ADRN1", "MES1"), min.cutoff = "q10", max.cutoff = "q90", blend = TRUE, cols = c("lightgrey", "orange", "navy"), combine = FALSE)

p[3] <- lapply(p[3], function(x) x+theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1)) + ylab("UMAP_1") + xlab("UMAP_2"))

p[4] <- lapply(p[4], function(x) x+theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1)) + ylab("MES") + xlab("ADRN"))

Fig2b <- p[3]
Fig2c <- p[4]

saveRDS(Fig2b, file = "Fig2b.rds")
saveRDS(Fig2c, file = "Fig2c.rds")

Fig2c
```
Adrenergic, mesenchymal and proliferation score calculater per cell.
Plotted in featurePlot
```{r Vlnplot of score}
#Plotting
VlnPlot(data.umap, features = c("ADRN1", "MES1"), group.by = "seurat_clusters", pt.size = 0, cols = cols_clust)
```
Plotted in VlnPlot for every cluster

```{r Heatmap of Adr/Mes genes}
#heatmap.data <- ScaleData(data.umap, features = c(adrn.genes2, mes.genes2))
#DoHeatmap(heatmap.data, features = c(adrn.genes2, mes.genes2), group.by = "seurat_clusters", label = FALSE)
```

```{r Again}
#heatmap.PT <- subset(heatmap.data, type == "PT")
#heatmap.LM <- subset(heatmap.data, type == "LM")
#DoHeatmap(heatmap.PT, features = c(adrn.genes2, mes.genes2), group.by = "seurat_clusters", label = FALSE)
```

```{r Again}
#DoHeatmap(heatmap.LM, features = c(adrn.genes2, mes.genes2), group.by = "seurat_clusters", label = FALSE)
#rm(heatmap.PT, heatmap.LM)
```
Heatmap av mes/adr/prolif gener i PT data
```{r Assigning Adr/Mes identity}
#Manually assigning Adrenergic, Mesenchymal and proliferative to clusters matching
data.umap$cellState <- "Unassigned"
data.umap$cellState[data.umap$seurat_clusters %in% c("1", "2", "3", "4", "5", "7", "8", "10")] <- "Adr"
data.umap$cellState[data.umap$seurat_clusters %in% c("6", "9")] <- "Mes"
#data.umap$cellState[data.umap$seurat_clusters %in% c("?")] <- "Proliferating"

data.umap$type_state <- paste(data.umap$type, data.umap$cellState, sep = "_")
data.umap$type_PDX_state <- paste(data.umap$PDX_type, data.umap$cellState, sep = "_")

saveRDS(data.umap, file = "SeuratObject_data.umap.rds")
data.umap <- readRDS("SeuratObject_data.umap.rds")

#Checking against reference genes again
#data.umap <- PrepSCTFindMarkers(data.umap)
#avg.expr <- AverageExpression(data.umap, assay = "SCT", return.seruat = FALSE)$SCT
#adrn.score <- colMeans(avg.expr[intersect(adrn.genes2, rownames(avg.expr)), ])
#mes.score <- colMeans(avg.expr[intersect(mes.genes2, rownames(avg.expr)), ])
#plot(adrn.score, mes.score, x.lab = "Adrenergic score", y.lab = "Mesenchymal score")
```
Plotting average expression of genes for cells and plotting against
```{r Dotplot + Scatterplot of identity}
#DotPlot(data.umap, features = c("PHOX2B","PRRX1","MKI67"), group.by = "cellState")
Fig2a <- FeatureScatter(data.umap, feature1 = "ADRN1", feature2 = "MES1", split.by = "seurat_clusters", group.by = "seurat_clusters", cols = cols_clust) + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1)) + ylab("MES") + xlab("ADRN")

saveRDS(Fig2a, file = "Fig2a.rds")
```
Sanity check
```{r UMAP of identity}
#DimPlot(data.umap, reduction = "umap_harmony", group.by = "cellState", label = TRUE, pt.size = 0.5) + NoLegend()
```
New UMAP with cellStates marked
```{r Barplot of cellstate per PDX}
onlyPTPDX1 <- subset(data.umap, PDX == "LU-NB-1" & type == "PT")
onlyPTPDX2 <- subset(data.umap, PDX == "LU-NB-2" & type == "PT")
onlyPTPDX3 <- subset(data.umap, PDX == "LU-NB-3" & type == "PT")
onlyLMPDX1 <- subset(data.umap, PDX == "LU-NB-1" & type == "LM")
onlyLMPDX2 <- subset(data.umap, PDX == "LU-NB-2" & type == "LM")
onlyLMPDX3 <- subset(data.umap, PDX == "LU-NB-3" & type == "LM")

CelllineType <- c(rep("PT LU-NB-1", 2), rep("LM LU-NB-1", 2), rep("PT LU-NB-2", 2), rep("LM LU-NB-2", 2), rep("PT LU-NB-3", 2), rep("LM LU-NB-3", 2))
Cellstate <- c(rep(c("MES", "ADRN"), 6))
Percentage <- c(c(100*length(WhichCells(onlyPTPDX1, expression = cellState == "Mes"))/length(colnames(onlyPTPDX1)),100* length(WhichCells(onlyPTPDX1, expression = cellState == "Adr"))/length(colnames(onlyPTPDX1))),
              c(100*length(WhichCells(onlyLMPDX1, expression = cellState == "Mes"))/length(colnames(onlyLMPDX1)), 100*length(WhichCells(onlyLMPDX1, expression = cellState == "Adr"))/length(colnames(onlyLMPDX1))),
              c(100*length(WhichCells(onlyPTPDX2, expression = cellState == "Mes"))/length(colnames(onlyPTPDX2)), 100*length(WhichCells(onlyPTPDX2, expression = cellState == "Adr"))/length(colnames(onlyPTPDX2))), 
              c(100*length(WhichCells(onlyLMPDX2, expression = cellState == "Mes"))/length(colnames(onlyLMPDX2)), 100*length(WhichCells(onlyLMPDX2, expression = cellState == "Adr"))/length(colnames(onlyLMPDX2))), 
              c(100*length(WhichCells(onlyPTPDX3, expression = cellState == "Mes"))/length(colnames(onlyPTPDX3)), 100*length(WhichCells(onlyPTPDX3, expression = cellState == "Adr"))/length(colnames(onlyPTPDX3))),
              c(100*length(WhichCells(onlyLMPDX3, expression = cellState == "Mes"))/length(colnames(onlyLMPDX3)), 100*length(WhichCells(onlyLMPDX3, expression = cellState == "Adr"))/length(colnames(onlyLMPDX3))))

barplotData <- data.frame(CelllineType, Cellstate, Percentage)
barplotData$Cellstate <- factor(barplotData$Cellstate, levels = c("MES", "ADRN"))
barplotData$CelllineType <- factor(barplotData$CelllineType, levels = c("PT LU-NB-1", "LM LU-NB-1", "PT LU-NB-2", "LM LU-NB-2", "PT LU-NB-3", "LM LU-NB-3"))
cols_state <- c("ADRN" = "orange", "MES" = "navy")

Fig2d <- ggplot(barplotData, aes(fill=Cellstate, y=Percentage, x=CelllineType, size=8)) + 
  geom_bar(position = "fill", stat = "identity", width = 0.9) + 
  ggtitle("Percentage ADRN/MES per PDX and Location") +
  scale_fill_manual(values = cols_state) +
  scale_y_continuous(labels = scales::label_percent(), expand = c(0, 0)) +
  scale_x_discrete(limits = c("PT LU-NB-1", "LM LU-NB-1", "", "PT LU-NB-2", "LM LU-NB-2", "", "PT LU-NB-3", "LM LU-NB-3"), labels = c("PT LU-NB-1", "LM LU-NB-1", "", "PT LU-NB-2", "LM LU-NB-2", "", "PT LU-NB-3", "LM LU-NB-3")) +
  #annotate(geom = "text", x = 1, y = -0.02, label = "PT", size = 3) +
  #annotate(geom = "text", x = 2, y = -0.02, label = "LM", size = 3) +
  #annotate(geom = "text", x = 4, y = -0.02, label = "PT", size = 3) +
  #annotate(geom = "text", x = 5, y = -0.02, label = "LM", size = 3) +
  #annotate(geom = "text", x = 7, y = -0.02, label = "PT", size = 3) +
  #annotate(geom = "text", x = 8, y = -0.02, label = "LM", size = 3) +
  theme_minimal() +
  geom_text(aes(label = ifelse(Percentage > 3, round(Percentage, 2), "")), 
            position = position_fill(vjust = 0.5), color = "white", size = 4) +
  theme(
  legend.position = "right",
  legend.title = element_blank(),
  axis.title.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.y = element_blank(),
  axis.line.y = element_line(color = "black", linewidth = 1.5, lineend = "square"),
  axis.ticks.y = element_line(color = "black", linewidth = 1.5),
  axis.ticks.length.y = unit(0.2, "cm"),
  axis.line.x = element_blank(),
  axis.text.x = element_text(angle = 45, hjust = 1, size = 15),
  axis.text.y = element_text(size = 15),
  panel.grid = element_blank(),
  text = element_text(family = "sans", size = 12, color = "black")
)

saveRDS(Fig2d, file = "Fig2d.rds")

rm(onlyPTPDX1, onlyPTPDX2, onlyPTPDX3, onlyLMPDX1, onlyLMPDX2, onlyLMPDX3, CelllineType, Percentage, Cellstate, barplotData)
```

##########################################################################################################
##Only LM data
```{r Filtering LM data}
#countMatrix <- GetAssayData(serLM, assay = "RNA", layer = "counts")
#countMatrix[is.na(countMatrix)] <- 0
#serLM <- SetAssayData(serLM, assay = "RNA", layer = "counts", new.data = countMatrix) #Remove NA's introduced #in decontX. 
#serLM <- subset(serLM, nFeature_RNA > 0) #Remove cells with no expressed genes (the expression was removed in #decontX as cells were decontamination)

##Showing Mitochondrial, Ribosomal, Hemoglobin, Platelate markers
#filteredLM <- serLM
#dim(filteredLM)
#table(filteredLM$mouse)
#table(filteredLM$PDX)

#filteredLM <- PercentageFeatureSet(filteredLM, pattern = "^MT-", col.name = "percent_mito")
#filteredLM <- PercentageFeatureSet(filteredLM, "^RP[SL]", col.name = "percent_ribo")
#filteredLM <- PercentageFeatureSet(filteredLM, "^HB[^(P|E|S)]", col.name = "percent_hb")
#filteredLM <- PercentageFeatureSet(filteredLM, "PECAM1|PF4", col.name = "percent_plat")

#Filter out cells and genes with low expression
#selectCell <- WhichCells(filteredLM, expression = nFeature_RNA > 100)
#selectFeat <- rownames(filteredLM)[Matrix::rowSums(filteredLM[["RNA"]]$counts) > 10]
#filteredLM <- subset(filteredLM, features = selectFeat, cells = selectCell)

##Filter out cells with high mitochondrial and low ribosomal expression
#filteredLM <- subset(filteredLM, percent_mito < 20 & percent_ribo > 5)

## Filter Mitochondrial, Ribosomal and Hemoglobin genes
#filteredLM <- filteredLM[!grepl("^MT-", rownames(filteredLM)), ]
#filteredLM <- filteredLM[!grepl("^RP[SL]", rownames(filteredLM)), ]
#filteredLM <- filteredLM[!grepl("^HB[^(P|E|S)]", rownames(filteredLM)), ]

##DoubletFinder
#LM.split <- SplitObject(filteredLM, split.by = "batch")
#LM.split <- lapply(LM.split, run_doubletfinder_custom)

#sglt_dblt_metadata <- data.frame(bind_rows(LM.split)) # merge to a single dataframe
#rownames(sglt_dblt_metadata) <- sglt_dblt_metadata$row_names # assign cell IDs to row names to ensure match
#sglt_dblt_metadata$row_names <- NULL
#filteredLM <- AddMetaData(filteredLM, sglt_dblt_metadata, col.name = 'doublet_finder')
#filteredLM <- filteredLM[, filteredLM@meta.data[, "doublet_finder"] == "Singlet"]

#rm(sglt_dblt_metadata, LM.split, selectCell, selectFeat)

#saveRDS(filteredLM, file = "SeuratObject_filteredLM.rds")
#filteredLM <- readRDS("SeuratObject_filteredLM.rds")

#dim(filteredLM)
#table(filteredLM$mouse)
#table(filteredLM$PDX)
```
Doing filtering with the same parameters
```{r Integrating and processing LM data}
#processedLM <- filteredLM
#processedLM[["RNA"]] <- split(processedLM[["RNA"]], f = processedLM$batch)

#processedLM <- NormalizeData(processedLM, verbose = FALSE)
#processedLM <- FindVariableFeatures(processedLM, verbose = FALSE)
#processedLM <- ScaleData(processedLM, verbose = FALSE)
#processedLM <- RunPCA(processedLM, verbose = FALSE)

#saveRDS(processedLM, file = "SeuratObject_processedLM.rds")
#processedLM <- readRDS("SeuratObject_processedLM.rds")

#Integrating
#integratedLM <- IntegrateLayers(object = processedLM, method = HarmonyIntegration, orig.reduction = "pca", #new.reduction = "harmony", verbose = FALSE)
#integratedLM[["RNA"]] <- JoinLayers(integratedLM[["RNA"]])

#saveRDS(integratedLM, file = "SeuratObject_integratedLM.rds")
#integratedLM <- readRDS("SeuratObject_integratedLM.rds")

#UMAP and tSNE
#finalLM <- RunUMAP(integratedLM, dims = 1:30, reduction = "harmony", reduction.name = "umap_harmony", verbose = 3FALSE)
#finalLMtSNE <- RunTSNE(integratedLM, reduction = "harmony", dims = 1:15, perplexity = 30, max_iter = 1000, theta = 0.5, eta = 200, num_threads = 0, verbose = FALSE)

#wrap_plots(
#  DimPlot(finalLM, reduction = "umap_harmony", group.by = "PDX", cols = cols_PDX_LM) + ggtitle("LM UMAP"),
#  DimPlot(finalLMtSNE, reduction = "tsne", group.by = "PDX", cols = cols_PDX_LM) + ggtitle("LM tSNE")
#)
```
Same processing and plotted in UMAP and tSNE
```{r Calculate clusters}
#finalLM <- FindNeighbors(finalLM, dims = 1:20, verbose = FALSE)
#finalLM <- FindClusters(finalLM, resolution = 1.2, verbose = FALSE)

#wrap_plots(
#  DimPlot(finalLM, reduction = "umap_harmony") + ggtitle("UMAP clustered"),
#  DimPlot(finalLM, reduction = "umap_harmony", group.by = "PDX", cols = cols_PDX_LM) + ggtitle("UMAP PDX"),
#  ncol = 2
#) + plot_layout(guides = "collect")
```
Plotting the clustered cells and based on PDX
```{r Adding Adr/Mes scores}
##Checking cellstates
#finalLM <- AddModuleScore(finalLM, features = list(adrn.genes2), name = "ADRN")
#finalLM <- AddModuleScore(finalLM, features = list(mes.genes2), name = "MES")
#finalLM <- AddModuleScore(finalLM, features = list(prolif.genes), name = "Prolif")
#Plotting
#FeaturePlot(finalLM, features = c("ADRN1", "MES1"), min.cutoff = "q10", max.cutoff = "q90", ncol = 3)
#VlnPlot(finalLM, features = c("ADRN1", "MES1"), group.by = "seurat_clusters", pt.size = 0)
```
Plotting cellstage markers in LM
```{r Heatmap of Adr/Mes genes}
#heatmap.finalLM <- ScaleData(finalLM, features = c(adrn.genes2, mes.genes2))
#DoHeatmap(heatmap.finalLM, features = c(adrn.genes2, mes.genes2), group.by = "seurat_clusters", label = FALSE)
#DoHeatmap(heatmap.finalLM, features = c(adrn.genes2, mes.genes2), group.by = "PDX", label = FALSE)
```
Heatmap av LM med Mes/Adr/prolif markers
```{r Assigning Adr/Mes identity}
##Manually assigning Adrenergic, Mesenchymal and proliferative to clusters matching
#finalLM$cellState <- "Unassigned"
#finalLM$cellState[finalLM$seurat_clusters %in% c("0", "1", "6", "3", "4", "5")] <- "Adrenergic"
#finalLM$cellState[finalLM$seurat_clusters %in% c("2", "7")] <- "Mesenchymal"

##Checking against reference genes again
#avg.exprLM <- AverageExpression(finalLM, assay = "RNA", return.seruat = FALSE)$RNA
#adrn.scoreLM <- colMeans(avg.exprLM[intersect(adrn.genes2, rownames(avg.exprLM)), ])
#mes.scoreLM <- colMeans(avg.exprLM[intersect(mes.genes2, rownames(avg.exprLM)), ])
#plot(adrn.scoreLM, mes.scoreLM, x.lab = "Adrenergic score", y.lab = "Mesenchymal score")

#DotPlot(finalLM, features = c("PHOX2B","PRRX1","MKI67"), group.by = "cellState")
#DimPlot(finalLM, reduction = "umap_harmony", group.by = "cellState", label = TRUE, pt.size = 0.5) + NoLegend()
```

```{r Heatmap of cluster genes}
#saveRDS(finalLM, file = "SeuratObject_finalLM.rds")
#finalLM <- readRDS("SeuratObject_finalLM.rds")
##Finding cluster markers and plotting in heatmap
#clusterMarkers <- FindAllMarkers(finalLM)
#clusterMarkers %>% 
#  group_by(cluster) %>%
#  dplyr::filter(p_val_adj < 0.05) %>%
#  slice_head(n=10) %>%
#  ungroup() -> top10Cluster
#clusterHeatmap <- ScaleData(finalLM, features = top10Cluster$gene)
#DoHeatmap(clusterHeatmap, features = top10Cluster$gene)
#rm(clusterHeatmap)
```
Top10 genes/cluster heatmap.
```{r DE Adrenergic vs Mesenchymal i LM data}
#Idents(finalLM) <- "cellState"
#AdrMesLM.markers <- FindMarkers(finalLM, ident.1 = "Adrenergic", ident.2 = "Mesenchymal", verbose = FALSE)
#uppGenesAdrMes <- AdrMesLM.markers$gene[which(AdrMesLM.markers$p_val_adj < 0.05 & AdrMesLM.markers$avg_log2FC > 0)]
#downGenesAdrMes <- AdrMesLM.markers$gene[which(AdrMesLM.markers$p_val_adj < 0.05 & AdrMesLM.markers$avg_log2FC < 0)]
#signGenesAdrMes <- c(uppGenesAdrMes, downGenesAdrMes)
#clusterHeatmap <- ScaleData(finalLM, features = signGenesAdrMes)
#DoHeatmap(clusterHeatmap, features = signGenesAdrMes, group.by = "PDX", label = FALSE)
#DoHeatmap(clusterHeatmap, features = signGenesAdrMes, group.by = "seurat_clusters", label = FALSE)
#rm(clusterHeatmap)
```

##########################################################################################################
##Differential expression analysis
Pseudobulking groups the cells together in the factors we want to compare. Gives more accurate analysis as it is less affected by outlier cells. 
```{r DE DESeq2 }
#pseudobulk
#data.pseudo <- AggregateExpression(data.umap, assays = "RNA", return.seurat = T, group.by = c("type", #"PDX_type"))

#Idents(data.pseudo) <- "type"

#data.de <- FindMarkers(object = data.pseudo, 
#                            ident.1 = "LM", 
#                            ident.2 = "PT",
#                            test.use = "DESeq2",
#                            min.pct = 0.25)

#saveRDS(data.de, file = "DEObject_Pseudobulk_LM-PT.rds")
data.de <- readRDS("DEObject_Pseudobulk_LM-PT.rds")

#rm(data.pseudo)
#gc()
#head(data.de, n = 60)
```
Showing first 60 differentially expressed genes.
```{r DESeq2 properly}
#bulked <- AggregateExpression(data.umap, assays(RNA), layer = "counts", group.by = c("type", "PDX", "mouse"), #return.seurat = FALSE)
#counts <- bulked$RNA
#colData <- data.frame(type_pdx = colnames(counts))
#
#colData$type <- sapply(strsplit(colData$type_pdx, "_"), `[`, 1)
#colData$type <- sapply(strsplit(colData$type_pdx, "_"), `[`, 2)
#DefaultAssay(data.umap) <- "RNA"
data.sce <- as.SingleCellExperiment(data.umap)
data.sce$model_type <- paste0(data.sce$mouse, "_", data.sce$type)
data.sce$group_id <- data.sce$type
data.sce$sample_id <- data.sce$orig.ident
data.sce$cluster_id <- "NB"

pbulked <- aggregateData(data.sce, assay = "counts", by = c("cluster_id", "model_type"), fun = "sum")
counts <- assay(pbulked, "NB")
colData <- as.data.frame(colData(pbulked))

colData$model_id <- colData$mouse
colData$model_id <- factor(colData$model_id)
colData$group_id <- factor(colData$group_id)

colData(pbulked)$model_id <- colData$model_id
colData(pbulked)$group_id <- colData$group_id

colData(pbulked)$model_id <- factor(colData(pbulked)$model_id)
colData(pbulked)$group_id <- factor(colData(pbulked)$group_id)

DeseqObj <- DESeqDataSetFromMatrix(countData = assay(pbulked), colData = colData(pbulked), design = ~ model_id + group_id)

DeseqObj <- DESeq(DeseqObj)
DeseqRes <- results(DeseqObj, contrast = c("group_id", "LM", "PT"))
DeseqRes <- DeseqRes[order(DeseqRes$padj),]
DeseqRes$gene <- rownames(DeseqRes)

saveRDS(DeseqRes, file = "DeseqRes_PTvsLM.rds")
DeseqRes <- readRDS("DeseqRes_PTvsLM.rds")

#DefaultAssay(data.umap) <- "SCT"

rm(data.sce, pbulked, colData, counts)
gc()
```

```{r Getting sign Gene list}
upGenesDES <- DeseqRes$gene[which(DeseqRes$padj < 0.01 & DeseqRes$log2FoldChange > 2)]
downGenesDES <- DeseqRes$gene[which(DeseqRes$padj < 0.01 & DeseqRes$log2FoldChange < -2)]
signGenesDES <- c(upGenesDES, downGenesDES)

saveRDS(upGenesDES, file = "List_upGenesDES_LM-PT.rds")
saveRDS(downGenesDES, file = "List_downGenesDES_LM-PT.rds")
saveRDS(signGenesDES, file = "List_signGenesDES_LM-PT.rds")

upGenesDES <- readRDS("List_upGenesDES_LM-PT.rds")
downGenesDES <- readRDS("List_downGenesDES_LM-PT.rds")
signGenesDES <- readRDS("List_signGenesDES_LM-PT.rds")
```


```{r Calculate significant genes}
##significant genes
#data.de$gene <- rownames(data.de)
#uppGenes <- data.de$gene[which(data.de$p_val_adj < 0.05 & data.de$avg_log2FC > 0.5)]
#downGenes <- data.de$gene[which(data.de$p_val_adj < 0.05 & data.de$avg_log2FC < -0.5)]
##signGenes <- data.de$gene[which(data.de$p_val_adj < 0.05)]
#signGenes <- c(uppGenes, downGenes)

#saveRDS(uppGenes, file = "List_uppGenes_LM-PT.rds")
#saveRDS(downGenes, file = "List_downGenes_LM-PT.rds")
#saveRDS(signGenes, file = "List_signGenes_LM-PT.rds")

uppGenes <- readRDS("List_uppGenes_LM-PT.rds")
downGenes <- readRDS("List_downGenes_LM-PT.rds")
signGenes <- readRDS("List_signGenes_LM-PT.rds")

#print(length(data.de$gene[which(data.de$p_val < 0.05)]))
#print(length(signGenes))
#print(length(uppGenes))
#print(length(downGenes))
```
Number of significantly expressed genes after setting cutoff at 0.05
```{r DE MAST Random intercepts (Crashes)}
#Using MAST and random intercepts
#library(MAST)
#exprMatr <- GetAssayData(data.umap, assay = "RNA", layer = "data")
#exprMatr <- as.matrix(exprMatr)
#cData <- data.umap@meta.data
#cData <- cData[, c("type", "PDX", "orig.ident", "nCount_RNA")]
#cData$type <- factor(cData$type)
#cData$PDX <- factor(cData$PDX)
#cData$orig.ident <- factor(cData$orig.ident)
#cData$log_umi <- log10(cData$nCount_RNA + 1)
#fData <- data.frame(gene_id <- rownames(exprMatr), row.names <- rownames(exprMatr))
#Forming sca object for MAST 
#scaMAST <- FromMatrix(exprsArray = exprMatr, cData = cData, fData = fData)

#zlm_fit <- zlm(formula = ~ type + log_umi + (1 | PDX),
#               sca = scaMAST,
#               method = "glmer",
#               ebayes = FALSE)
#summary_fit <- summary(zlm_fit, doLRT = "typeLM")
#lrt <- summary_fit$datatable
#de_resultsMAST <- lrt[lrt$component == "H" & lrt$contrast == "typeLM", ]
#de_resultsMAST$FDR <- p.adjust(de_resultsMAST$`Pr(>Chisq)`, method = "fdr")
#de_resultsMAST <- de_resultsMAST[order(de_results$FDR), ]
#head(de_resultMast, n = 40)

#Skapa lista av significanta gener
```

```{r DE DESeq2 Adr vs AdrLM and Mes vs Mes LM (not needed)}
#Doing once for adrenergic and once for mesenchymal
#pseudo_type <- AggregateExpression(data.umap, assays = "RNA", return.seurat = TRUE, group.by = c("type_state", "PDX_type", #"type_mouse"))
#Idents(pseudo_type) <- "type_state"


#Adr.de <- FindMarkers(object = pseudo_type,
#                      ident.1 = "LM-Adrenergic",
#                      ident.2 = "PT-Adrenergic",
#                      test.use = "DESeq2")
#Mes.de <- FindMarkers(object = pseudo_type,
#                      ident.1 = "LM-Mesenchymal",
#                      ident.2 = "PT-Mesenchymal",
#                      test.use = "DESeq2")
#head(Adr.de, n = 20)
#head(Mes.de, n = 20)

#significant genes Adr
#Adr.de$gene <- rownames(Adr.de)
#uppGenesAdr <- Adr.de$gene[which(Adr.de$p_val_adj < 0.05 & Adr.de$avg_log2FC > 0)]
#downGenesAdr <- Adr.de$gene[which(Adr.de$p_val_adj < 0.05 & Adr.de$avg_log2FC < 0)]
#signGenesAdr <- c(uppGenesAdr, downGenesAdr)
#print(length(Adr.de$gene[which(Adr.de$p_val < 0.05)]))
#print(length(signGenesAdr))
#print(length(uppGenesAdr))
#print(length(downGenesAdr))

#significant genes Mes
#Mes.de$gene <- rownames(Mes.de)
#uppGenesMes <- Mes.de$gene[which(Mes.de$p_val_adj < 0.05 & Mes.de$avg_log2FC > 0)]
#downGenesMes <- Mes.de$gene[which(Mes.de$p_val_adj < 0.05 & Mes.de$avg_log2FC < 0)]
#signGenesMes <- c(uppGenesMes, downGenesMes)
#print(length(Mes.de$gene[which(Mes.de$p_val < 0.05)]))
#print(length(signGenesMes))
#print(length(uppGenesMes))
#print(length(downGenesMes))

```

##Visulising significant genes
```{r Vlnplot of some genes}
# generate violin plot
data.umap$type <- factor(x = data.umap$type, levels = c("PT", "LM"))
Idents(data.umap) <- "type"
#VlnPlot(data.umap, features = signGenes[1:8], idents = c("PT", "LM"), group.by = "type", ncol = 4) 
#VlnPlot(data.umap, features = signGenes[1:8], idents = c("PT", "LM"), group.by = "PDX_type", ncol = 4)
```
Showing expression levels in 8 most significant genes in all cells, grouped by tumor type, and mouse. 
```{r Featureplot of some genes}
#Featureplot
#FeaturePlot(data.umap, features = signGenes[1:4])
#FeaturePlot(data.umap, features = signGenes[5:8])
#FeaturePlot(data.umap, features = signGenes[9:12])
```
Showing cells expression of significant genes in UMAP projection 
```{r Dotplot sign genes}
Fig3b <- DotPlot(data.umap, features = signGenesDES, group.by = "PDX_type") + RotatedAxis()
Fig3b

saveRDS(Fig3b, file = "Fig3b.rds")
```

```{r Heatmap sign genes}
#data.doheat <- ScaleData(data.umap, features = signGenesDES)
#DoHeatmap(data.doheat, features = signGenesDES, group.by = "type", group.colors = cols_type)
#DoHeatmap(data.doheat, features = signGenesDES, group.by = "PDX_type", group.colors = cols_PDX_type, label = #FALSE)
#DoMultiBarHeatmap(data.doheat, features = signGenes[40], layer = "scale.data", group.by = "type", additional.group.by = "PDX")
#rm(data.doheat)
```
Shows heatmap of most significant genes. (Show less or more genes?)
```{r Volcanoplot}
#Volcanoplot
volcplot.de <- data.de %>%
  mutate(gene = rownames(data.de),
         significance = case_when(
           avg_log2FC > 2 & p_val_adj < 0.05 ~ "Upregulated",
           avg_log2FC < -2 & p_val_adj < 0.05 ~ "Downregulated",
           TRUE ~ "Not significant"
         ))
top_genes <- volcplot.de %>%
  arrange(p_val_adj) %>%
  dplyr::slice(1:20)

ggplot(volcplot.de, aes(x = avg_log2FC, y = -log10(p_val_adj), color = significance)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("Downregulated" = "#6BAED6", #"#00AFBB"
                                "Not significant" = "grey",
                                "Upregulated" = "#E6550D")) +
  geom_vline(xintercept = c(-2, 2), linetype = "dashed") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  geom_point(size = 1) +
  geom_text_repel(data = top_genes, aes(label = gene), size = 2, max.overlaps = 20) +
  theme_minimal() +
  labs(title = "PT-LM: Differentially Expressed Genes",
       x = "Average log2 Fold Change",
       y = "-log10 Adjusted P-value")
rm(volcplot.de, top_genes)
```
Volcanoplot showing expression against significance, cutoffs at 0.5(SD) and 0.05 significant. Top 20 most significant genes labeled.
```{r Volcanoplot new gene list}
#Volcanoplot
#volcplot.de <- DeseqRes %>%
#  mutate(gene = rownames(DeseqRes),
#         significance = case_when(
#           log2FoldChange > 0.5 & padj < 0.05 ~ "Upregulated",
#           log2FoldChange < -0.5 & padj < 0.05 ~ "Downregulated",
#           TRUE ~ "Not significant"
#         ))
Volcano.Deseq <- data.frame(gene = rownames(DeseqRes), log2FoldChange = DeseqRes$log2FoldChange, padj = DeseqRes$padj, significance = ifelse(DeseqRes$log2FoldChange > 2 & DeseqRes$padj < 0.01, "Upregulated", ifelse(DeseqRes$log2FoldChange < -2 & DeseqRes$padj < 0.01, "Downregulated", "Not significant")))

top_genes <- Volcano.Deseq %>%
  arrange(padj) %>%
  dplyr::slice(1:20)

Fig3a <- ggplot(Volcano.Deseq, aes(x = log2FoldChange, y = -log10(padj), color = significance)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("Downregulated" = "#6BAED6", #"#00AFBB"
                                "Not significant" = "grey",
                                "Upregulated" = "#E6550D")) +
  geom_vline(xintercept = c(-2, 2), linetype = "dashed") +
  geom_hline(yintercept = -log10(0.01), linetype = "dashed") +
  geom_point(size = 1) +
  geom_text_repel(data = top_genes, aes(label = gene), size = 2, max.overlaps = 20) +
  theme_minimal() +
  labs(title = "PT-LM: Differentially Expressed Genes",
       x = "Average log2 Fold Change",
       y = "-log10 Adjusted P-value")

Fig3a

saveRDS(Fig3a, file = "Fig3a.rds")

rm(volcplot.de, top_genes)

```

```{r clustered heatmap sign genes}
#data.heatmap <- subset(data.umap, features = signGenesDES)

#pseudo$cluster <- unlist(lapply(split(pseudo, pseudo[["PDX_type"]]), function(df) {
#  n <- nrow(df)
#  paste0(df[["PDX_type"]][1], "_", ceiling(seq_len(n) / 100))}))
#data.heatmap$pseudocluster <- pseudo$cluster

##Create annotation dataframes
#avg_exp <- AggregateExpression(data.heatmap, group.by = "pseudocluster", assays = "RNA", slot = "data")$RNA
#annotations_df <- data.frame(CellType = sub("-\\d+$", "", colnames(avg_exp)), row.names = colnames(avg_exp))
#splitcelltype <- strsplit(annotations_df$CellType, " ")
#type <- sapply(splitcelltype, "[", 1)
#PDX <- sapply(splitcelltype, "[", 2)
#annotations_df$CellType <- type
#annotations_df$CellLine <- PDX



#genes_ann <- rownames(avg_exp)
#gene_ann_df <- data.frame(Regulation = ifelse(genes_ann %in% uppGenes, "LM", ifelse(genes_ann %in% downGenes, "PT", "Neutral")), row.names = genes_ann, stringsAsFactors = FALSE)
#data.heatmap <- as.matrix(data.heatmap@assays$RNA@layers$data) #used before adding AggregateExpression
#col_fun = colorRamp2(c(-0.5, 1, 2.5), c("#FF00FF", "#000000", "#FFFF00"))
#annotations_col = list(CellType = cols_type, CellLine = cols_PDX, Regulation = c(PT = "#6BAED6", LM = "#E6550D"))
#col_2 <- colorRampPalette(c("#FF00FF", "black", "yellow"))(101)

#col_fun(seq(-0.25, 2.25, length.out = 101)), , breaks = seq(-0, 2, length.out = 101)

#pheatmap(avg_exp, scale = "row", breaks = seq(-1, 2.5, length.out = 101), annotation_row = gene_ann_df, #annotation_col = annotations_df, show_colnames = F, color = col_2, annotation_colors = annotations_col, #fontsize_row = 6, legend_breaks = c(0, 1, 2), legend_labels = c("0", "1", "2"), annotation_names_row = FALSE, #annotation_names_col = FALSE)
#rm(data.heatmap, pseudo, avg_exp, annotations_df, col_fun, annotations_col)
```
Clustered heatmap showing how celltype correspond to generic clustering. Grouped inside type, for 100 cells in each.
```{r Clustered heatmap of DE genes with ComplexHeatmap}
clusterHeatmap <- ScaleData(data.umap, features = signGenesDES)
heatmapSer <- subset(clusterHeatmap, features = signGenesDES)
heatmapMatr <- GetAssayData(heatmapSer, assay = "RNA", layer = "scale.data")
heatmapMatr <- heatmapMatr[order(factor(rownames(heatmapMatr), levels = signGenesDES)), ]

options(expressions = 500000) #Because it is very very nested, gives error
#dend <- as.dendrogram(hclust(dist(t(heatmapMatr))))
#dend <- collapse_branch(dend, tol = 0.5)
#dend %>% collapse_branch(tol = 20) %>% ladderize %>% hang.dendrogram(hang = 0)
#dend <- cut(dend, h = 10)
#plot(dend, main = "please")

heatmapAnn <- HeatmapAnnotation(Cluster = heatmapSer$seurat_clusters, PDX = heatmapSer$PDX, Location = heatmapSer$type, col = list(Cluster = cols_clust, PDX = cols_PDX, Location = cols_type), annotation_label = c("Cluster", "PDX", "Type"), show_legend = FALSE)
heatmapAnnR <- rowAnnotation(Signature = ifelse(rownames(heatmapMatr) %in% upGenesDES, "LM", "PT"), show_legend = FALSE, col = list(Signature = cols_type), annotation_label = c(" "))


h <- Heatmap(heatmapMatr, name = "clusterHeatmap", cluster_rows = TRUE, cluster_columns = TRUE, show_column_dend = TRUE, show_column_names = FALSE, show_row_dend = TRUE , top_annotation = heatmapAnn, left_annotation = heatmapAnnR, col = col_fun, use_raster = FALSE, column_title = "Clustered heatmap", show_heatmap_legend = FALSE, row_names_gp = gpar(fontsize = 6))

matrixLGND <- Legend(col_fun = col_fun, title = "Expression")
clusterLGND <- Legend(at = 0:10, title = "Clusters", legend_gp = gpar(fill = cols_clust))
cellLineLGND <- Legend(labels = c("LU-NB-1", "LU-NB-2", "LU-NB-3"), title = "PDX", legend_gp = gpar(fill = cols_PDX))
typeLGND <- Legend(labels = c("PT", "LM"), title = "Tumor type", legend_gp = gpar(fill = cols_type))
GeneSignLGND <- Legend(labels = c("PT", "LM"), title = "Gene signature", legend_gp = gpar(fill = cols_type))
LGNDs <- packLegend(matrixLGND, clusterLGND, cellLineLGND, typeLGND, GeneSignLGND)

Fig3e <- draw(h, annotation_legend_list = LGNDs)

saveRDS(Fig3e, file = "Fig3e.rds")
Fig3e <- readRDS("Fig3e.rds")

rm(h, heatmapMatr, heatmapAnn, heatmapAnnR, matrixLGND, clusterLGND, cellLineLGND, GeneSignLGND, LGNDs, heatmapSer, clusterHeatmap)
```


```{r Corrplot}
data_mat <- GetAssayData(data.umap, assay = "RNA", layer = "data")
data_mat <- data_mat[signGenesDES, , drop = FALSE]
data_matt <- t(as.matrix(data_mat))
corrProb <- cor(data_matt, method = "spearman", use = "complete.obs") # creates matrix of p-values for correlation between the significant genes

corrProbAnn <- HeatmapAnnotation(Signature = ifelse(colnames(corrProb) %in% upGenesDES, "LM", "PT"), show_legend = FALSE, col = list(Signature = cols_type), annotation_label = c(" "))
corrProbAnnR <- rowAnnotation(Signature = ifelse(rownames(corrProb) %in% upGenesDES, "LM", "PT"), show_legend = TRUE, col = list(Signature = cols_type), annotation_label = c(" "))
corr_col <- colorRamp2(c(-0.2, 0.15, 0.6), c("blue", "white", "red"))

h <- Heatmap(corrProb, name = "Correlation", clustering_distance_rows = "spearman", clustering_distance_columns = "spearman", cluster_rows = TRUE, cluster_columns = TRUE, show_column_dend = TRUE, col = corr_col, show_column_names = TRUE, show_row_dend = TRUE , top_annotation = corrProbAnn, left_annotation = corrProbAnnR, show_heatmap_legend = TRUE, row_names_side = "right", column_dend_side = "top", column_names_rot = 45, row_gap = unit(1, "mm"), column_gap = unit(1, "mm"), row_dend_width = unit(2, "cm"), column_dend_height = unit(2, "cm"), row_names_gp = gpar(fontsize = 6), column_names_gp = gpar(fontsize = 6))

Fig3d <- draw(h)

saveRDS(Fig3d, file = "Fig3d.rds")


```

Clustering genes correlation and plotting in heatmap
###########################################################################################################
##Gene set enrichment analysis
Visualizing biological pathways uppregulated
```{r GO enrichment}
GO_enrich_Up <- enrichGO(gene = upGenesDES, OrgDb = 'org.Hs.eg.db', ont = "BP", keyType = "SYMBOL")
GO_enrich_Down <- enrichGO(gene = downGenesDES, OrgDb = 'org.Hs.eg.db', ont = "BP", keyType = "SYMBOL")
GO_Bar_Up <- barplot(GO_enrich_Up, drop = TRUE, showCategory = 20, title = "GO Upregulated", font.size = 8)
GO_Bar_Down <- barplot(GO_enrich_Down, drop = TRUE, showCategory = 20, title = "GO Downregulated", font.size = 8)

(GO_Bar_Up | GO_Bar_Down) +  plot_layout(guides = "collect")
```

Gene ontology enrichment analysis (Biological process)
```{r Grouped GO enrichment}
simplified_GO_enrich_Down <- simplify(GO_enrich_Down, cutoff = 0.7, by = "p.adjust", select_fun = min, measure = "Wang")
barplot(simplified_GO_enrich_Down, drop = TRUE, showCategory = 10, title = "GO Downregulated Grouped")
```
Grouping pathways

```{r KEGG enrichment}
#KEGG pathway enrichment analysis
uppGenesEntrez <- mapIds(org.Hs.eg.db, keys = upGenesDES, column = "ENTREZID", keytype = "SYMBOL", multiVals = "first")
downGenesEntrez <- mapIds(org.Hs.eg.db, keys = downGenesDES, column = "ENTREZID", keytype = "SYMBOL", multiVals = "first")
signGenesEntrez <- mapIds(org.Hs.eg.db, keys = signGenesDES, column = "ENTREZID", keytype = "SYMBOL", multiVals = "first")
KEGG_enrich_Up <- enrichKEGG(gene = uppGenesEntrez, organism = 'human', pvalueCutoff = 0.05, keyType = "ncbi-geneid")
KEGG_enrich_Down <- enrichKEGG(gene = downGenesEntrez, organism = 'human', pvalueCutoff = 0.05, keyType = "ncbi-geneid")
KEGG_Bar_Up <- barplot(KEGG_enrich_Up, drop = TRUE, showCategory = 20, title = "KEGG Upregulated", font.size = 8)
KEGG_Bar_Down <- barplot(KEGG_enrich_Down, drop = TRUE, showCategory = 20, title = "KEGG Downregulated", font.size = 8)
(KEGG_Bar_Up | KEGG_Bar_Down) +  plot_layout(guides = "collect")
```
KEGG enrichment analysis (Pathways)

```{r NCG enrichment (no results)}
NCG_enrich_Up <- enrichNCG(gene = uppGenesEntrez, pvalueCutoff = 0.05, pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500)
NCG_enrich_Down <- enrichNCG(gene = downGenesEntrez, pvalueCutoff = 0.05, pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500)
NCG_Bar_Down <- barplot(NCG_enrich_Down, drop = TRUE, showCategory = 20, title = "NCG Enrichment", font.size = 8)
NCG_Bar_Up <- barplot(NCG_enrich_Up, drop = TRUE, showCategory = 20, title = "NCG Enrichment", font.size = 8)
(NCG_Bar_Up | NCG_Bar_Down) +  plot_layout(guides = "collect")
```
Network of cancer gene enrichment analysis(comparing to a library of known cancer genes and their expression and function)

```{r KEGG GSEA}
geneList <- DeseqRes$log2FoldChange # skapa lista med gener och deras värden
names(geneList) <- DeseqRes$gene
gene.df <- bitr(names(geneList), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db) #översätter till entrez

geneList <- geneList[gene.df$SYMBOL]
names(geneList) <- gene.df$ENTREZID #ny lista med samma värden men entrez id istället
geneList <- sort(geneList, decreasing = TRUE) #sorterar listan

gseKegg <- gseKEGG(geneList = geneList, organism = "hsa", pvalueCutoff = 0.05, verbose = FALSE) #enkel gseKEGG
gse_Plot_KEGG <- dotplot(gseKegg, showCategory = 10, split=".sign", title = "GSEA KEGG pathways")
```
gseKEGG, väger in värdena, inte bara lista av signifikanta gener.

```{r GO GSEA dotplot}
gseGo <- gseGO(geneList = geneList, OrgDb = org.Hs.eg.db, keyType = "ENTREZID", ont = "BP", pvalueCutoff = 0.05, verbose = FALSE)
gse_Plot_GO <- dotplot(gseGo, showCategory = 10, split=".sign", title = "GSEA GO pathways")
(gse_Plot_GO | gse_Plot_KEGG) +  plot_layout(guides = "collect")
```
gseGO, kombination, gse fast med GO library
```{r Same but grouped}
simplified_gseGO <- simplify(gseGo, cutoff = 0.7, by = "p.adjust", select_fun = min, measure = "Wang")
simplified_Plot_gseGO <- dotplot(simplified_gseGO, showCategory = 10, split=".sign", title = "GSEA GO pathways grouped")
(simplified_Plot_gseGO | gse_Plot_KEGG) +  plot_layout(guides = "collect")
```
With simplified

```{r KEGG and GO GSEA Barplot}
dfGO <- gseGo@result
dfGO <- dfGO[dfGO$p.adjust < 0.05, ]
# Top 10 up and down
df_mainGO <- rbind(head(dfGO[order(-dfGO$NES), ], 10), head(dfGO[order(dfGO$NES), ], 10))
df_mainGO$Direction <- ifelse(df_mainGO$NES > 0, "Up", "Down")

dfKEGG <- gseKegg@result
dfKEGG <- dfKEGG[dfKEGG$p.adjust < 0.05, ]
# Top 10 up and down
df_mainKEGG <- rbind(head(dfKEGG[order(-dfKEGG$NES), ], 10), head(dfKEGG[order(dfKEGG$NES), ], 10))
df_mainKEGG$Direction <- ifelse(df_mainKEGG$NES > 0, "Up", "Down")

NES_Bar_GO <- ggplot(df_mainGO, aes(x = NES, y = reorder(Description, NES), fill = p.adjust)) +
  geom_col(width = 0.8) +
  labs(x = "Normalized Enrichment Score (NES)", y = NULL, title = "GO GSEA") +
  scale_fill_gradient(low  = "red", high = "blue", name = "p.adjust") +
  theme_minimal(base_size = 12) +
  theme(legend.title = element_blank(), plot.title = element_text(face = "bold")
  )
NES_Bar_KEGG <- ggplot(df_mainKEGG, aes(x = NES, y = reorder(Description, NES), fill = p.adjust)) +
  geom_col(width = 0.8) +
  labs(x = "Normalized Enrichment Score (NES)", y = NULL, title = "KEGG GSEA") +
  scale_fill_gradient(low  = "red", high = "blue", name = "p.adjust") +
  theme_minimal(base_size = 12) +
  theme(legend.title = element_blank(), plot.title = element_text(face = "bold")
  )

(NES_Bar_GO | NES_Bar_KEGG) +  plot_layout(guides = "collect")
Fig4a <- NES_Bar_KEGG

saveRDS(Fig4a, file = "Fig4a.rds")
```

```{r Barplot with p.adjust as length}
dfKEGG <- gseKegg@result
dfKEGG <- dfKEGG[dfKEGG$p.adjust < 0.05, ]
# Top 10 up and down
df_mainKEGG <- head(dfKEGG[order(dfKEGG$p.adjust), ], 100)
df_mainKEGG$Direction <- ifelse(df_mainKEGG$NES > 0, "Up", "Down")

NES_Bar_KEGG <- ggplot(df_mainKEGG, aes(x = -log(p.adjust), y = reorder(Description, -p.adjust), fill = NES)) +
  geom_col(width = 0.8) +
  labs(x = "-log 10 Adjusted P-value", y = NULL, title = "KEGG GSEA") +
  scale_fill_gradient(low  = "blue", high = "red", name = "NES") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold")
  )
NES_Bar_KEGG

NES_Bar_KEGG <- ggplot(df_mainKEGG, aes(x = reorder(Description, p.adjust), y = -log(p.adjust), fill = NES)) +
  geom_col(width = 0.8) +
  labs(x = NULL, y = "-log 10 Adjusted P-value", title = "KEGG GSEA") +
  scale_fill_gradient(low  = "blue", high = "red", name = "NES") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), 
        axis.text.x = element_text(angle = 70, hjust = 1)
  )
NES_Bar_KEGG
```

Barplot instead of dotplot, less variables, but still NES and p.adjust
```{r Again but grouped}
dfGOsimp <- simplified_gseGO@result
dfGOsimp <- dfGOsimp[dfGOsimp$p.adjust < 0.05, ]
# Top 10 up and down
df_mainGOsimp <- rbind(head(dfGOsimp[order(-dfGOsimp$NES), ], 10), head(dfGOsimp[order(dfGOsimp$NES), ], 10))
df_mainGOsimp$Direction <- ifelse(df_mainGOsimp$NES > 0, "Up", "Down")

simplified_NES_Bar_GO <- ggplot(df_mainGOsimp, aes(x = NES, y = reorder(Description, NES), fill = p.adjust)) +
  geom_col(width = 0.8) +
  labs(x = "Normalized Enrichment Score (NES)", y = NULL, title = "GO GSEA grouped") +
  scale_fill_gradient(low  = "red", high = "blue", name = "p.adjust") +
  theme_minimal(base_size = 12) +
  theme(legend.title = element_blank(), plot.title = element_text(face = "bold")
  )
(simplified_NES_Bar_GO | NES_Bar_KEGG) +  plot_layout(guides = "collect")
```

```{r Pathview (stemcell signaling one)}
library(pathview)

# Plot specific KEGG pathways (with fold change) 
## pathway.id : KEGG pathway identifier
pathview(gene.data = geneList, pathway.id = "04550", species = "human") 
```
Can plot specific pathways and foldchange for genes involved.
```{r GO enrichment cellular compartment (no result)}
#Checking which cellular compartment the corresponding protein is aktive in
compartmentGOUp <- enrichGO(gene = upGenesDES, OrgDb = 'org.Hs.eg.db', ont = "CC", keyType = "SYMBOL")
compartmentGODown <- enrichGO(gene = downGenesDES, OrgDb = 'org.Hs.eg.db', ont = "CC", keyType = "SYMBOL")
barplot(compartmentGOUp, drop = TRUE, showCategory = 20, title = "GO Upregulated Cellular Compartment", font.size = 8)
barplot(compartmentGODown, drop = TRUE, showCategory = 20, title = "GO Downregulated Cellular Compartment", font.size = 8)
```
Ett sätt att testa vilka cellular components våra gener är involverade i
```{r GO cellular compartment get.GO.for.genes (not working yet)}
#get.GO.for.genes(signGenes, bio = T, cel = T, mol = F)
#get.GO.for.genes(signGenesAdr, bio = T, cel = T, mol = F)
#get.GO.for.genes(signGenesMes, bio = T, cel = T, mol = F)

```



#################################################################################################
##Identifying genes of interest
Looking for genes in good pathways and genes in many pathways
```{r List of sign genes in most pathways + fold change}
simplified_gseGOX <- setReadable(simplified_gseGO, 'org.Hs.eg.db', 'ENTREZID')
gseGoX <- setReadable(gseGo, 'org.Hs.eg.db', 'ENTREZID')
gseKeggX <- setReadable(gseKegg, 'org.Hs.eg.db', 'ENTREZID')

##GO
le_GO_df <- gseGoX@result %>%
  select(ID, Description, core_enrichment) %>%
  filter(!is.na(core_enrichment)) %>%
  tidyr::separate_rows(core_enrichment, sep = "/") %>%
  rename(Gene = core_enrichment)

geneListSign <- geneList[names(geneList) %in% signGenesEntrez]
#Also gene list with symbol, not entrez
geneListSymbol <- DeseqRes$log2FoldChange # skapa lista med gener och deras värden
names(geneListSymbol) <- DeseqRes$gene
geneListSymbol <- geneListSymbol[gene.df$SYMBOL]
names(geneListSymbol) <- gene.df$SYMBOL

gene_freq_GO <- sort(table(le_GO_df$Gene), decreasing = TRUE)
gene_freq_GO <- as.data.frame(gene_freq_GO)
colnames(gene_freq_GO) <- c("Gene", "Pathway count")
rank_df <- tibble(Gene = names(geneListSymbol), rank_metric = as.numeric(geneListSymbol))
gene_freq_GO <- gene_freq_GO %>% left_join(rank_df, by = "Gene")

gene_freq_GO <- gene_freq_GO %>% arrange(desc(`Pathway count`), desc(abs(rank_metric)))

#gene_rank_GO <- intersect(names(geneList), le_GO_df$Gene)
#geneList[gene_rank_GO]

gene_freq_GO_sign <- subset(gene_freq_GO, Gene %in% signGenesDES)
gene_freq_GO_sign

rm(gene_freq_GO, gene_freq_GO_sign, le_GO_df)

##GO simplified
le_GO_df <- simplified_gseGOX@result %>%
  select(ID, Description, core_enrichment) %>%
  filter(!is.na(core_enrichment)) %>%
  tidyr::separate_rows(core_enrichment, sep = "/") %>%
  rename(Gene = core_enrichment)

gene_freq_GO <- sort(table(le_GO_df$Gene), decreasing = TRUE)
gene_freq_GO <- as.data.frame(gene_freq_GO)
colnames(gene_freq_GO) <- c("Gene", "Pathway count")
gene_freq_GO <- gene_freq_GO %>% left_join(rank_df, by = "Gene")

gene_freq_GO <- gene_freq_GO %>% arrange(desc(`Pathway count`), desc(abs(rank_metric)))

gene_freq_GO_sign <- subset(gene_freq_GO, Gene %in% signGenesDES)
gene_freq_GO_sign

#KEGG
le_KEGG_df <- gseKeggX@result %>%
  select(ID, Description, core_enrichment) %>%
  filter(!is.na(core_enrichment)) %>%
  tidyr::separate_rows(core_enrichment, sep = "/") %>%
  rename(Gene = core_enrichment)

gene_freq_KEGG <- sort(table(le_KEGG_df$Gene), decreasing = TRUE)
gene_freq_KEGG <- as.data.frame(gene_freq_KEGG)
colnames(gene_freq_KEGG) <- c("Gene", "Pathway count")
gene_freq_KEGG <- gene_freq_KEGG %>% left_join(rank_df, by = "Gene")

gene_freq_KEGG <- gene_freq_KEGG %>% arrange(desc(`Pathway count`), desc(abs(rank_metric)))

gene_freq_KEGG_sign <- subset(gene_freq_KEGG, Gene %in% signGenesDES)
gene_freq_KEGG_sign
```
Lista över gener, antal pathways inkluderad i och foldChange(rank_metric)
```{r pathway plot}
gseaplot2(
  gseGo,
  geneSetID = "GO:0006119",
  title = "Oxidative phosphorylation"
)
```
Plot över generna i en pathway och var de är i vår lista


```{r Heatplot of pathways and their genes}
#heatplot(gseGoX, showCategory = 10, foldChange = geneList)
heatplot(simplified_gseGOX, showCategory = 10, foldChange = geneList)
heatplot(gseKeggX, showCategory = 10, foldChange = geneList)
#heatplot(gseGoX, showCategory = 10, foldChange = geneListSign)
heatplot(simplified_gseGOX, showCategory = 10, foldChange = geneListSign)
heatplot(gseKeggX, showCategory = 10, foldChange = geneListSign)

heatplot(gseKeggX, showCategory = c("Signaling pathways regulating pluripotency of stem cells", "Hedgehog signaling pathway", "TGF-beta signaling pathway", "Axon guidance", "Wnt signaling pathway"), foldChange = geneListSign)
#heatplot(simplified_gseGOX, showCategory = c("negative chemotaxis", "ephrin receptor signaling pathway", "axon guidance", "establishment or maintenance of apical/basal cell polarity", "regulation of small GTPase mediated signal transduction"), foldChange = geneListSign)
heatplot(gseKeggX, showCategory = c("Signaling pathways regulating pluripotency of stem cells", "TGF-beta signaling pathway"), foldChange = geneList) + scale_fill_gradient2(low = "blue", mid = "white", high = "red", name = "Fold Change")

heatplot(gseKeggX, showCategory = c("Cytokine-cytokine receptor interaction", "Pathways in cancer", "PI3K-Akt signaling pathway", "NOD-like receptor signaling pathway", "TNF signaling pathway", "Integrin signaling"), foldChange = geneList) + theme(axis.text.x = element_text(size = 4, lineheight = 0.8), axis.text.y = element_text(size = 6)) + scale_fill_gradient2(low = "blue", mid = "white", high = "red", name = "Fold Change")

Fig4b <- heatplot(gseKeggX, showCategory = c("Cytokine-cytokine receptor interaction", "NOD-like receptor signaling pathway"), foldChange = geneList) + theme(axis.text.x = element_text(size = 8, lineheight = 0.8), axis.text.y = element_text(size = 10)) + scale_fill_gradient2(low = "blue", mid = "white", high = "red", name = "Fold Change")

Fig4b

saveRDS(Fig4b, file = "Fig4b.rds")
Fig4b <- readRDS("Fig4b.rds")
```
Heatmap showing what genes are in a Pathway and which of these are significant, as well as how strong the change is


Extra dotplot, same as heatplot.
```{r Interconnected pathways and their genes}
#Top 20 pathways
#cnetplot(
#  gseGoX,
#  showCategory = 20,
#  foldChange = geneListSign,
#  node_label = "all",
#  fc_threshold = 0.3
#)
cnetplot(
  simplified_gseGOX,
  showCategory = 40,
  foldChange = geneListSign,
  node_label = "all",
  fc_threshold = 0.3
)
Fig4c <- cnetplot(
  gseKeggX,
  showCategory = 81,
  foldChange = geneListSign,
  node_label = "all",
  fc_threshold = 0.1
)+ scale_colour_gradientn(values = c(-2, -1, 2), colours = c("blue", "white", "red"))

saveRDS(Fig4b, file = "Fig4b.rds")
Fig4b <- readRDS("Fig4b.rds")

Fig4c

```

################################ PT och LM scores
Visualisera vidare

```{r Vlnplots of all significant genes}
data.umap$PDX_type <- factor(x = data.umap$PDX_type, levels = c("PT LU-NB-1", "PT LU-NB-2", "PT LU-NB-3", "LM LU-NB-1", "LM LU-NB-2", "LM LU-NB-3"))
data.umap$type_state <- factor(x = data.umap$type_state, levels = c("PT_Adr", "PT_Mes", "LM_Adr", "LM_Mes"))

VlnPlot(data.umap, features = signGenesDES, group.by = "seurat_clusters", pt.size = 0, stack = TRUE, flip = TRUE, cols = c(rep("#E6550D", 126), rep("#6BAED6", 55))) +
        theme(legend.position = "none")+ NoLegend()

DotPlot(data.umap, features = signGenesDES, group.by = "seurat_clusters")

saveRDS(Fig3c, file = "Fig3c.rds")

#VlnPlot(onlyPT, features = upGenesDES, group.by = "seurat_clusters", pt.size = 0, stack = TRUE, flip = TRUE, #cols = c(rep("#E6550D", 126))) + theme(legend.position = "none")
```
Vlnplot av alla gener i celllinjer och Adrenerga/Mesenchymala


```{r Adding LM and PT score + UMAP and Vlnplots}
#PT signature and LM signature
#Calculate score for each cell
data.umap <- AddModuleScore(data.umap, features = list(upGenesDES), name = "LMScore")
data.umap <- AddModuleScore(data.umap, features = list(downGenesDES), name = "PTScore")

#Featureplot of score
#FeaturePlot(data.umap, features = c("LMScore1", "PTScore1"), min.cutoff = "q10", max.cutoff = "q90", ncol = 3)
#Vinplot on cluster/cellstate
VlnPlot(data.umap, features = c("LMScore1", "PTScore1"), group.by = "type", pt.size = 0)
VlnPlot(data.umap, features = c("LMScore1", "PTScore1"), group.by = "PDX_type", pt.size = 0)
VlnPlot(data.umap, features = c("LMScore1", "PTScore1"), group.by = "type_state", pt.size = 0)
#VlnPlot(data.umap, features = c("LMScore1", "PTScore1"), group.by = "seurat_clusters", pt.size = 0)

#FeatureScatter(data.umap, feature1 = "PTScore1", feature2 = "LMScore1", split.by = "PDX", group.by = "type", cols = cols_type)

#VlnPlot(onlyPT, features = c("LMScore1", "PTScore1"), group.by = "seurat_clusters", pt.size = 0)
```
LM och PT score visualiserat
```{r find cut-off}
onlyLM <- subset(data.umap, type == "LM")
LMScoreMean <- mean(onlyLM$LMScore1)
LMScoreSD <- sd(onlyLM$LMScore1)
LMcutoff <- LMScoreMean - LMScoreSD

#cells <- c(colnames(onlyPT), colnames(onlyLM))
#Location <- c(rep("PT", length(colnames(onlyPT))), rep("LM", length(colnames(onlyLM))))
#Values <- data.umap$LMScore1
#Values <- as.character(Values)

#LMscoreDF <- data.frame(cells, Location, Values)

```
Finding 1SD from LM data. For LM profile cut-off

```{r Setting cut-off and assigning identity. + Barplot of percantages}
#LMcells <- WhichCells(data.umap, expression = LMScore1 > LMcutoff)
#PTcells <- WhichCells(data.umap, expression = PTScore1 > 0.1)
#data.umap$signature <- ifelse(colnames(data.umap) %in% LMcells & colnames(data.umap) %in% #PTcells, "Both", ifelse(colnames(data.umap) %in% LMcells, "LM", ifelse(colnames(data.umap) %in% #PTcells, "PT", "None")))
onlyPT <- subset(data.umap, type == "PT")
onlyPT$LMprofile <- ifelse(onlyPT$LMScore1 > LMcutoff, "LM-like", "not LM-like")
onlyPT1 <- subset(onlyPT, PDX == "LU-NB-1")
onlyPT2 <- subset(onlyPT, PDX == "LU-NB-2")
onlyPT3 <- subset(onlyPT, PDX == "LU-NB-3")
onlyLM <- subset(data.umap, type == "LM")

saveRDS(data.umap, file = "SeuratObject_data.umap_withScores.rds")
data.umap <- readRDS("SeuratObject_data.umap_withScores.rds")

saveRDS(onlyPT, file = "SeuratObject_onlyPT_withScores.rds")
onlyPT <- readRDS("SeuratObject_onlyPT_withScores.rds")

#DimPlot(onlyPT, group.by = "signature", cols = c(LM = "#E6550D", PT = "#6BAED6", Both = "#8B00FF", None = "#D3D3D3"), order = c("Both", "LM", "PT", "None")) + ggtitle("PT cells by signature")


Fig5b <- DimPlot(onlyPT1, group.by = "LMprofile", cols = c("LM-like" = "#E6550D", "not LM-like" = "#D3D3D3"), order = c("LM-like", "not LM-like"), pt.size = 1.2) + NoLegend() + ggtitle("PT LU-NB-1") + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 12, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1),legend.position = "inside", legend.position.inside = c(0.7, 0.7), legend.title=element_text(size=4)) + ylab("UMAP_2") + xlab("UMAP_1")
Fig5c <- DimPlot(onlyPT2, group.by = "LMprofile", cols = c("LM-like" = "#E6550D", "not LM-like" = "#D3D3D3"), order = c("LM-like", "not LM-like"), pt.size = 1.2) + NoLegend() + ggtitle("PT LU-NB-2") + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 12, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1),legend.position = "inside", legend.position.inside = c(0.7, 0.7), legend.title=element_text(size=4)) + ylab("UMAP_2") + xlab("UMAP_1")
Fig5d <- DimPlot(onlyPT3, group.by = "LMprofile", cols = c("LM-like" = "#E6550D", "not LM-like" = "#D3D3D3"), order = c("LM-like", "not LM-like"), pt.size = 1.2) + NoLegend() + ggtitle("PT LU-NB-3") + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 12, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1),legend.position = "inside", legend.position.inside = c(0.7, 0.7), legend.title=element_text(size=4)) + ylab("UMAP_2") + xlab("UMAP_1")

Fig5b
Fig5c
Fig5d

saveRDS(Fig5b, file = "Fig5b.rds")
saveRDS(Fig5c, file = "Fig5c.rds")
saveRDS(Fig5d, file = "Fig5d.rds")

Fig5b <- readRDS("Fig5b.rds")
Fig5c <- readRDS("Fig5c.rds")
Fig5d <- readRDS("Fig5d.rds")

wrap_plots(Fig5b, Fig5c, Fig5d, ncol = 3)

#FeatureScatter(data.umap, feature1 = "PTScore1", feature2 = "LMScore1", split.by = "PDX", group.by = "signature", cols = c(LM = "#E6550D", PT = "#6BAED6", Both = "#8B00FF", None = "#D3D3D3"))

#FeatureScatter(onlyPT, feature1 = "ADRN1", feature2 = "MES1", split.by = "PDX", group.by = "LMprofile", cols = c("LM+" = "#E6550D", "LM-" = "#D3D3D3"))

#FeatureScatter(onlyPT, feature1 = "ADRN1", feature2 = "MES1", split.by = "seurat_clusters", group.by = "LMprofile", cols = c("LM+" = "#E6550D", "LM-" = "#D3D3D3"))

#VlnPlot(onlyPT, features = c("LMScore1", "PTScore1"), group.by = "PDX", split.by = "cellState", pt.size = 0)
```

```{r Barplot, percentage of signatures}
CellLine <- c("PT LU-NB-1", "PT LU-NB-2", "PT LU-NB-3", "Extra")
Signature <- c(rep(c("LM"), 3), "Extra")
Percentage <- c(c(length(WhichCells(onlyPT1, expression = LMprofile == "LM-like"))/length(colnames(onlyPT1))), c(length(WhichCells(onlyPT2, expression = LMprofile == "LM-like"))/length(colnames(onlyPT2))), c(length(WhichCells(onlyPT3, expression = LMprofile == "LM-like"))/length(colnames(onlyPT3))), c(0.015))

barplotData <- data.frame(CellLine, Signature, Percentage)

Fig5a <- ggplot(barplotData, aes(fill=Signature, y=Percentage, x=CellLine)) + 
  geom_col() + 
  ggtitle("Percentage cells with LM profile") +
  scale_fill_manual(values = c("LM" = "#E6550D", "Extra" = "white")) +
  scale_y_continuous(breaks = seq(0, 0.015, by = 0.005), labels = scales::label_percent(), expand = c(0, 0)) +
  scale_x_discrete(limits = c("PT LU-NB-1", "PT LU-NB-2", "PT LU-NB-3", "Extra"), labels = c("PT LU-NB-1", "PT LU-NB-2", "PT LU-NB-3", "")) +
  theme_minimal() +
  geom_text(aes(label = 100*round(Percentage, 4)), position = position_dodge(width = 1), vjust = -0.5, color = "black", size = 4) +
  theme(
  legend.position = "right",
  legend.title = element_blank(),
  axis.title.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.y = element_blank(),
  axis.line.y = element_line(color = "black", linewidth = 1.5, lineend = "square"),
  axis.ticks.y = element_line(color = "black", linewidth = 1.5),
  axis.ticks.length.y = unit(0.2, "cm"),
  axis.line.x = element_blank(),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 15),
  panel.grid = element_blank(),
  text = element_text(family = "sans", size = 12, color = "black")
)

Fig5a

saveRDS(Fig5a, file = "Fig5a.rds")

Fig5a <- readRDS("Fig5a.rds")
```



```{r DESeq2 with DESeq2 package, LM+ PT cells}
PT.sce <- as.SingleCellExperiment(onlyPT)
PT.sce$model_type <- paste0(PT.sce$mouse, "_", PT.sce$LMprofile)
PT.sce$group_id <- PT.sce$LMprofile
PT.sce$sample_id <- PT.sce$orig.ident
PT.sce$cluster_id <- "NB"

PTpbulked <- aggregateData(PT.sce, assay = "counts", by = c("cluster_id", "model_type"), fun = "sum")
PTcounts <- assay(PTpbulked)
PTcolData <- as.data.frame(colData(PTpbulked))

PTcolData$model_id <- PTcolData$mouse
PTcolData$model_id <- factor(PTcolData$model_id)
PTcolData$group_id <- factor(PTcolData$group_id)

colData(PTpbulked)$model_id <- PTcolData$model_id
colData(PTpbulked)$group_id <- PTcolData$group_id

colData(PTpbulked)$model_id <- factor(colData(PTpbulked)$model_id)
colData(PTpbulked)$group_id <- factor(colData(PTpbulked)$group_id)

PTDeseqObj <- DESeqDataSetFromMatrix(countData = assay(PTpbulked), colData = colData(PTpbulked), design = ~ model_id + group_id)

PTDeseqObj <- DESeq(PTDeseqObj)
PTDeseqRes <- results(PTDeseqObj, contrast = c("group_id", "LM-like", "not LM-like"))
PTDeseqRes <- PTDeseqRes[order(PTDeseqRes$padj),]
PTDeseqRes$gene <- rownames(PTDeseqRes)

saveRDS(PTDeseqRes, file = "PTDeseqRes_PTvsLM.rds")

PTDeseqRes <- readRDS("PTDeseqRes_PTvsLM.rds")

```

```{r Getting sign Genes from DESeq2}
PTupGenesDES <- PTDeseqRes$gene[which(PTDeseqRes$padj < 0.01 & PTDeseqRes$log2FoldChange > 2)]
PTdownGenesDES <- PTDeseqRes$gene[which(PTDeseqRes$padj < 0.01 & PTDeseqRes$log2FoldChange < -2)]
PTsignGenesDES <- c(PTupGenesDES, PTdownGenesDES)

saveRDS(PTupGenesDES, file = "List_PTupGenesDES_LM-PT.rds")
saveRDS(PTdownGenesDES, file = "List_PTdownGenesDES_LM-PT.rds")
saveRDS(PTsignGenesDES, file = "List_PTsignGenesDES_LM-PT.rds")

PTupGenesDES <- readRDS("List_PTupGenesDES_LM-PT.rds")
PTdownGenesDES <- readRDS("List_PTdownGenesDES_LM-PT.rds")
PTsignGenesDES <- readRDS("List_PTsignGenesDES_LM-PT.rds")
```

```{r}
sharedLMandLMlike <- intersect(PTupGenesDES, upGenesDES)
# Q: What kind of plot is good to show that some genes in upGenes are differentially expressed in LM-like cells in PT already?
# A: Venn diagram? Barplot of how many genes in upGenes are also in PTupGenes? Heatmap of expression of shared genes in PT and LM?
# Q: Maybe a Venn diagram, how can this be done in R?
# A: Using the VennDiagram package, you can create a Venn diagram to show the overlap between the two gene sets. Here's an example of how to do this:
library(VennDiagram)
venn.plot <- venn.diagram(
  x = list("LM-DEG" = upGenesDES, "LM-like-DEG" = PTupGenesDES),
  category.names = c("LM-DEG", "LM-like-DEG"),
  filename = NULL,
  fill = c("#A61C3C", "#E6550D"),
  alpha = 0.6,
  col = "black",
  lwd = 2,
  cex = 2,
  fontface = "bold",
  cat.cex = 2,
  cat.fontface = "bold",
  cat.col = c("#A61C3C", "#E6550D"),
  cat.pos = c(-20, 20),
  cat.dist = c(0.05, 0.05),
  margin = 0.1,
  output=TRUE
)
venn.plot

install.packages("UpSetR")
library(UpSetR)
data <- list("LM-DEG" = upGenesDES, "LM-like-DEG" = PTupGenesDES)
upset(fromList(data))

#Create a plot with ggplot, with each gene as a dot, on a plane with padj from LM-DEG on x-axis and padj from LM-like-DEG on y-axis, and color the dots by whether they are in both sets or only one set, using ggplot2.
#First create a data frame with the genes, their padj from both sets, and whether they are in both sets or only one set.
LMandLMlikeGenes <- union(upGenesDES, PTupGenesDES) 

df <- data.frame(gene = LMandLMlikeGenes, padj_1 = DeseqRes$padj[match(LMandLMlikeGenes, DeseqRes$gene)], padj_2 = PTDeseqRes$padj[match(LMandLMlikeGenes, PTDeseqRes$gene)])
# Define threshold
threshold <- 0.01
# Add significance groups (Venn categories)
df <- df %>%
  mutate(
    group = case_when(
      padj_1 < threshold & padj_2 < threshold ~ "Both",
      padj_1 < threshold & padj_2 >= threshold ~ "LM DEG",
      padj_1 >= threshold & padj_2 < threshold ~ "LM-like DEG",
      TRUE ~ "Neither"
    ),
    logp1 = -log10(padj_1),
    logp2 = -log10(padj_2)
  )
# Plot
ggplot(df, aes(x = logp1, y = logp2, color = group)) +
  geom_point(size = 2, alpha = 0.8) +
  #Add labels for shared significant genes above padj 8 for oneof the sets
  geom_text_repel(data = subset(df, group == "Both" & (logp1 > 8 | logp2 > 8)), aes(label = gene), size = 3, max.overlaps = 20) +
  scale_y_continuous(limits = c(0, 30), expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(limits = c(0, 40), expand = expansion(mult = c(0, 0.1))) +
  # Threshold lines
  geom_vline(xintercept = -log10(threshold), linetype = "dashed") +
  geom_hline(yintercept = -log10(threshold), linetype = "dashed") +
  # Colors (match your red theme)
  scale_color_manual(values = c("Both" = "#7F0000", "LM DEG" = "#D55E00", "LM-like DEG" = "#A61C3C")) +
  labs(x = expression(-log[10](padj)~" LM DEG"),
    y = expression(-log[10](padj)~" LM-like DEG"),
    color = "Significance") +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  )
```

```{r Volcplot from DESeq2}
#Volcanoplot
PTVolcano.Deseq <- data.frame(gene = rownames(PTDeseqRes), log2FoldChange = PTDeseqRes$log2FoldChange, padj = PTDeseqRes$padj, significance = ifelse(PTDeseqRes$log2FoldChange > 2 & PTDeseqRes$padj < 0.01, "Upregulated", ifelse(PTDeseqRes$log2FoldChange < -2 & PTDeseqRes$padj < 0.01, "Downregulated", "Not significant")))

PTtop_genes <- PTVolcano.Deseq %>%
  arrange(padj) %>%
  dplyr::slice(1:20)

Fig5e <- ggplot(PTVolcano.Deseq, aes(x = log2FoldChange, y = -log10(padj), color = significance)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("Downregulated" = "#6BAED6", #"#00AFBB"
                                "Not significant" = "grey",
                                "Upregulated" = "#E6550D")) +
  geom_vline(xintercept = c(-2, 2), linetype = "dashed") +
  geom_hline(yintercept = -log10(0.01), linetype = "dashed") +
  geom_point(size = 1) +
  geom_text_repel(data = PTtop_genes, aes(label = gene), size = 2, max.overlaps = 20) +
  theme_minimal() +
  labs(title = "LM-like: Differentially Expressed Genes",
       x = "Average log2 Fold Change",
       y = "-log10 Adjusted P-value")

Fig5e

saveRDS(Fig5e, "Fig5e.rds")
Fig5e <- readRDS("Fig5e.rds")

rm(volcplot.de, top_genes)
```

```{r gseKEGG of LM signature PT cells}
LMscoreGeneList <- PTDeseqRes$log2FoldChange # skapa lista med gener och deras värden
names(LMscoreGeneList) <- PTDeseqRes$gene
LMscoreGene.df <- bitr(names(LMscoreGeneList), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db) #översätter till entrez

LMscoreGeneList <- LMscoreGeneList[LMscoreGene.df$SYMBOL]
names(LMscoreGeneList) <- LMscoreGene.df$ENTREZID #ny lista med samma värden men entrez id istället
LMscoreGeneList <- sort(LMscoreGeneList, decreasing = TRUE) #sorterar listan

LMscoreGseKegg <- gseKEGG(geneList = LMscoreGeneList, organism = "hsa", pvalueCutoff = 0.05, verbose = FALSE) #enkel gseKEGG
LMscoreGseGoFrst <- gseGO(geneList = LMscoreGeneList, OrgDb = org.Hs.eg.db, keyType = "ENTREZID", ont = "BP", pvalueCutoff = 0.05, verbose = FALSE)
LMscoreGseGo <- simplify(LMscoreGseGoFrst, cutoff = 0.7, by = "p.adjust", select_fun = min, measure = "Wang")

LMdfGO <- LMscoreGseGo@result
LMdfGO <- LMdfGO[LMdfGO$p.adjust < 0.05, ]
# Top 10 up and down
LMdf_mainGO <- rbind(head(LMdfGO[order(-LMdfGO$NES), ], 10), head(LMdfGO[order(LMdfGO$NES), ], 10))
LMdf_mainGO$Direction <- ifelse(LMdf_mainGO$NES > 0, "Up", "Down")

LMdfKEGG <- LMscoreGseKegg@result
LMdfKEGG <- LMdfKEGG[LMdfKEGG$p.adjust < 0.05, ]
# Top 10 up and down
LMdf_mainKEGG <- rbind(head(LMdfKEGG[order(-LMdfKEGG$NES), ], 10), head(LMdfKEGG[order(LMdfKEGG$NES), ], 10))
LMdf_mainKEGG$Direction <- ifelse(LMdf_mainKEGG$NES > 0, "Up", "Down")

LMscoreNES_Bar_GO <- ggplot(LMdf_mainGO, aes(x = NES, y = reorder(Description, NES), fill = p.adjust)) +
  geom_col(width = 0.8) +
  labs(x = "Normalized Enrichment Score (NES)", y = NULL, title = "GO GSEA LMscore") +
  scale_fill_gradient(low  = "red", high = "blue", name = "p.adjust") +
  theme_minimal(base_size = 12) +
  theme(legend.title = element_blank(), plot.title = element_text(face = "bold")
  )
LMscoreNES_Bar_KEGG <- ggplot(LMdf_mainKEGG, aes(x = NES, y = reorder(Description, NES), fill = p.adjust)) +
  geom_col(width = 0.8) +
  labs(x = "Normalized Enrichment Score (NES)", y = NULL, title = "KEGG GSEA LMscore") +
  scale_fill_gradient(low  = "red", high = "blue", name = "p.adjust") +
  theme_minimal(base_size = 12) +
  theme(legend.title = element_blank(), plot.title = element_text(face = "bold")
  )

(LMscoreNES_Bar_GO | LMscoreNES_Bar_KEGG) +  plot_layout(guides = "collect")
LMscoreNES_Bar_KEGG


```

```{r Pathways on p.adjust}
LMdfKEGG <- LMscoreGseKegg@result
LMdfKEGG <- LMdfKEGG[LMdfKEGG$p.adjust < 0.05, ]
# Top 10 up and down
LMdf_mainKEGG <- head(LMdfKEGG[order(LMdfKEGG$p.adjust), ], 10)
LMdf_mainKEGG$Direction <- ifelse(LMdf_mainKEGG$NES > 0, "Up", "Down")

LM_p_Bar_KEGG <- ggplot(LMdf_mainKEGG, aes(x = -log(p.adjust), y = reorder(Description, -p.adjust), fill = NES)) +
  geom_col(width = 0.8) +
  labs(x = "-log 10 Adjusted P-value", y = NULL, title = "KEGG GSEA") +
  scale_fill_gradient(low  = "#FFD4D4", high = "red", name = "NES") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold")
  )
Fig5f <- LM_p_Bar_KEGG

saveRDS(Fig5f, file = "Fig5f.rds")

Fig5f
```
Pathways från LM score
```{r cnetPlots of pathways}
LMscoreGseKeggX <- setReadable(LMscoreGseKegg, 'org.Hs.eg.db', 'ENTREZID')
LMscoreGseGoX <- setReadable(LMscoreGseGo, 'org.Hs.eg.db', 'ENTREZID')

cnetplot(
  LMscoreGseKeggX,
  showCategory = 10,
  foldChange = LMscoreGeneList,
  node_label = "all",
  fc_threshold = 0.5
)
cnetplot(
  LMscoreGseGoX,
  showCategory = 10,
  foldChange = LMscoreGeneList,
  node_label = "all",
  fc_threshold = 0.5
)
```

```{r Trying to find trajectory in umap with singature marked}
#wrap_plots(
#  DimPlot(onlyPT, reduction = "pca", group.by = "signature", dims = 1:2),
#  DimPlot(onlyPT, reduction = "pca", group.by = "signature", dims = 3:4),
#  DimPlot(onlyPT, reduction = "pca", group.by = "signature", dims = 5:6),
#  DimPlot(onlyPT, reduction = "pca", group.by = "signature", dims = 7:8),
#  DimPlot(onlyPT, reduction = "pca", group.by = "signature", dims = 9:10),
#  DimPlot(onlyPT, reduction = "pca", group.by = "signature", dims = 11:12),
#  DimPlot(onlyPT, reduction = "pca", group.by = "signature", dims = 13:14),
#  DimPlot(onlyPT, reduction = "pca", group.by = "signature", dims = 15:16),
#  DimPlot(onlyPT, reduction = "pca", group.by = "signature", dims = 17:18),
#  ncol = 3
#) + plot_layout(guides = "collect")
```

```{r Plots with BM signature as well}
upGenesBM <- readRDS("List_upGenesBM_BM-PT.rds")
downGenesBM <- readRDS("List_downGenesBM_BM-PT.rds")
signGenesBM <- readRDS("List_signGenesBM_BM-PT.rds")
DeseqResBM <- readRDS("DeseqResBM_BMvsPT.rds")
BMcells <- readRDS("List_BMposLMcells.rds")

#Genes from BM need to be present in both datasets
BMupGenes <- intersect(upGenesBM, rownames(data.umap))
BMdownGenes <- intersect(downGenesBM, rownames(data.umap))
BMsignGenes <- intersect(signGenesBM, rownames(data.umap))

#Combining and intersecting
CommonUpGenes <- intersect(BMupGenes, upGenesDES)
CommonDownGenes <- intersect(BMdownGenes, downGenesDES)

CombinedUpList <- union(BMupGenes, upGenesDES)
CombinedGeneList <- union(BMsignGenes, signGenesDES)

LMupBMdown <- intersect(upGenesDES, downGenesBM)
BMupLMdown <- intersect(upGenesBM, downGenesDES)

onlyPT <- AddModuleScore(onlyPT, features = list(BMupGenes), name = "BMScore")
onlyPT$BMprofileSymb <- ifelse(colnames(onlyPT) %in% BMcells, "BM+", "BM-")


FeatureScatter(onlyPT, feature1 = "LMScore1", feature2 = "BMScore1", split.by = "PDX", group.by = "PDX", cols = cols_PDX)
Fig6a <- FeatureScatter(onlyPT, feature1 = "LMScore1", feature2 = "BMScore1", split.by = "seurat_clusters", group.by = "seurat_clusters", cols = cols_clust, ncol = 5) + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1)) + ylab("BMScore") + xlab("LMScore")

saveRDS(Fig6a, "Fig6a.rds")
Fig6a <- readRDS("Fig6a.rds")
Fig6a

DimPlot(onlyPT, split.by = "PDX", group.by = "LMprofileSymb", cols = c("LM+" = "#E6550D", "LM-" = "#D3D3D3"), order = c("LM+", "LM-"), pt.size = 1.2) + NoLegend() + ggtitle("PT LU-NB-3") + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 12, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1),legend.position = "inside", legend.position.inside = c(0.8, 0.8)) + ylab("UMAP_2") + xlab("UMAP_1")
DimPlot(onlyPT, split.by = "PDX", group.by = "BMprofileSymb", cols = c("BM+" = "yellow", "BM-" = "#D3D3D3"), order = c("LM+", "LM-"), pt.size = 1.2) + NoLegend() + ggtitle("PT LU-NB-3") + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 12, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1),legend.position = "inside", legend.position.inside = c(0.8, 0.8)) + ylab("UMAP_2") + xlab("UMAP_1")

DimPlot(onlyPT, split.by = "PDX", group.by = "profile", cols = c("BM+" = "yellow", "LM+" = "#E6550D", "Both" = "orange", "None" = "#D3D3D3"), order = c("Both", "LM+", "BM+", "None"), pt.size = 1.2) + NoLegend() + ggtitle("PT LU-NB-3") + theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 12, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1),legend.position = "inside", legend.position.inside = c(0.8, 0.8)) + ylab("UMAP_2") + xlab("UMAP_1")

saveRDS(Fig6b, "Fig6b.rds")
Fig6b <- readRDS("Fig6b.rds")

Fig6b
```

```{r Corrplot between genes}
data_mat <- GetAssayData(data.umap, assay = "RNA", layer = "data")
data_mat <- data_mat[CombinedUpList, , drop = FALSE]
data_matt <- t(as.matrix(data_mat))
corrProb <- cor(data_matt, method = "spearman", use = "complete.obs")

corrProbAnn <- HeatmapAnnotation(Signature = ifelse(colnames(corrProb) %in% upGenesDES & colnames(corrProb) %in% upGenesBM, "Both", ifelse(colnames(corrProb) %in% upGenesDES & !(colnames(corrProb) %in% upGenesBM), "LM", "BM")), show_legend = FALSE, col = list(Signature = c("Both" = "orange", "LM" = "red", "BM" = "yellow")), annotation_label = c(" "))
corrProbAnnR <- rowAnnotation(Signature = ifelse(colnames(corrProb) %in% upGenesDES & colnames(corrProb) %in% upGenesBM, "Both", ifelse(colnames(corrProb) %in% upGenesDES & !(colnames(corrProb) %in% upGenesBM), "LM", "BM")), show_legend = TRUE, col = list(Signature = c("Both" = "orange", "LM" = "red", "BM" = "yellow")), annotation_label = c(" "))
corr_col <- colorRamp2(c(-0.2, 0.15, 0.6), c("blue", "white", "red"))

h <- Heatmap(corrProb, name = "Correlation", clustering_distance_rows = "spearman", clustering_distance_columns = "spearman", cluster_rows = TRUE, cluster_columns = TRUE, show_column_dend = TRUE, col = corr_col, show_column_names = TRUE, show_row_dend = TRUE , top_annotation = corrProbAnn, left_annotation = corrProbAnnR, show_heatmap_legend = TRUE, row_names_side = "right", column_dend_side = "top", column_names_rot = 45, row_gap = unit(1, "mm"), column_gap = unit(1, "mm"), row_dend_width = unit(2, "cm"), column_dend_height = unit(2, "cm"), row_names_gp = gpar(fontsize = 5), column_names_gp = gpar(fontsize = 5))

Fig6c <- draw(h)

saveRDS(Fig6c, "Fig6c.rds")
Fig6c <- readRDS("Fig6c.rds")

draw(h)
```

```{r FoldChange BMvsPT genes}
BMDeseqRes <- DeseqResBM[rownames(DeseqResBM) %in% rownames(DeseqRes), ]
LMDeseqRes <- DeseqRes[rownames(DeseqRes) %in% rownames(BMDeseqRes), ]

PTLMandBMFC <- data.frame(gene = rownames(LMDeseqRes), log2FoldChangeLM = LMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), LMDeseqRes$gene)], log2FoldChangeBM = BMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), BMDeseqRes$gene)], significance = ifelse(LMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), LMDeseqRes$gene)] > 2 & BMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), BMDeseqRes$gene)] > 2, "High Both", ifelse(LMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), LMDeseqRes$gene)] < -2 & BMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), BMDeseqRes$gene)] < -2, "Low Both", ifelse(LMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), LMDeseqRes$gene)] < -2 & BMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), BMDeseqRes$gene)] > 2, "High BM, Low LM", ifelse(LMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), LMDeseqRes$gene)] > 2 & BMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), BMDeseqRes$gene)] < -2, "High LM, Low BM", ifelse(LMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), LMDeseqRes$gene)] < 2 & BMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), BMDeseqRes$gene)] > 2, "High BM only", ifelse(LMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), LMDeseqRes$gene)] > 2 & BMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), BMDeseqRes$gene)] < 2, "High LM only", ifelse(LMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), LMDeseqRes$gene)] < -2, "Low LM only", ifelse(BMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), BMDeseqRes$gene)] < -2, "Low BM only", "Not significant")))))))))

AllHighlightedGenes <- PTLMandBMFC[which(PTLMandBMFC$significance == "High Both" | PTLMandBMFC$significance == "Low Both"| PTLMandBMFC$significance == "High BM, Low LM"| PTLMandBMFC$significance == "High LM, Low BM"| PTLMandBMFC$significance == "High BM only"| PTLMandBMFC$significance == "High LM only"| PTLMandBMFC$significance == "Low BM only"| PTLMandBMFC$significance == "Low LM only"), ]

Fig6d <- ggplot(PTLMandBMFC, aes(x = log2FoldChangeLM, y = log2FoldChangeBM, color = significance)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("Low Both" = "#6BAED6", #"#00AFBB"
                                "Not significant" = "grey",
                                "High Both" = "orange",
                                "High BM only" = "gold",
                                "High LM only" = "red",
                                "High LM, Low BM" = "purple",
                                "High BM, Low LM" = "forestgreen",
                                "Low LM only" = "#088F8F",
                                "Low BM only" = "#4e4e94"
                                )) +
  geom_vline(xintercept = c(-2, 2), linetype = "dashed") +
  geom_hline(yintercept = c(-2, 2), linetype = "dashed") +
  geom_point(size = 1) +
  geom_text_repel(data = AllHighlightedGenes, aes(label = gene), size = 2, max.overlaps = 20) +
  theme_minimal() +
  labs(title = "Fold change of LM and BM genes",
       x = "Average log2 Fold Change LM-PT",
       y = "Average log2 Fold Change BM-PT")

saveRDS(Fig6d, "Fig6d.rds")
Fig6d <- readRDS("Fig6d.rds")

Fig6d
```

```{r same but p.adjust}
PTLMandBMPA <- data.frame(gene = rownames(LMDeseqRes), p.adjustLM = LMDeseqRes$padj, p.adjustBM = BMDeseqRes$padj, significance = ifelse(LMDeseqRes$padj < 0.01 & BMDeseqRes$padj < 0.01, "Upregulated", "Not significant"))

AllHighlightedGenes <- PTLMandBMPA[which(PTLMandBMPA$significance == "Upregulated"), ]

Fig6g <- ggplot(PTLMandBMPA, aes(x = -log10(p.adjustLM), y = -log10(p.adjustBM), color = significance)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("Not significant" = "grey",
                                "Upregulated" = "#E6550D")) +
  geom_vline(xintercept = c(-log10(0.01)), linetype = "dashed") +
  geom_hline(yintercept = c(-log10(0.01)), linetype = "dashed") +
  geom_point(size = 1) +
  geom_text_repel(data = AllHighlightedGenes, aes(label = gene), size = 2, max.overlaps = 20) +
  theme_minimal() +
  labs(title = "Fold change of LM and BM genes",
       x = "Adjusted P-value LM-PT",
       y = "Adjusted P-value BM-PT")

saveRDS(Fig6g, "Fig6g.rds")
Fig6g <- readRDS("Fig6g.rds")

Fig6g
```

```{r again concordance plot}
df <- data.frame(gene = CombinedUpList, padj_1 = DeseqRes$padj[match(CombinedUpList, DeseqRes$gene)], padj_2 = DeseqResBM$padj[match(CombinedUpList, DeseqResBM$gene)])
df <- df %>%
  mutate(
    padj_1 = ifelse(is.na(padj_1), 1, padj_1),
    padj_2 = ifelse(is.na(padj_2), 1, padj_2)
  )
# Define threshold
threshold <- 0.01
# Add significance groups (Venn categories)
df <- df %>%
  mutate(
    group = case_when(
      padj_1 < threshold & padj_2 < threshold ~ "Both",
      padj_1 < threshold & padj_2 >= threshold ~ "LM DEG",
      padj_1 >= threshold & padj_2 < threshold ~ "BM DEG",
      TRUE ~ "Neither"
    ),
    logp1 = -log10(padj_1),
    logp2 = -log10(padj_2)
  )
# Plot
ggplot(df, aes(x = logp1, y = logp2, color = group)) +
  geom_point(size = 2, alpha = 0.8) +
  #Add labels for shared significant genes above padj 8 for oneof the sets
  geom_text_repel(data = subset(df, group == "Both"  | (logp1 > 20 | logp2 > 5)), aes(label = gene), size = 3, max.overlaps = 20) +
  scale_y_continuous(limits = c(-0.1, 10), expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(limits = c(-0.1, 80), expand = expansion(mult = c(0, 0.1))) +
  # Threshold lines
  geom_vline(xintercept = -log10(threshold), linetype = "dashed") +
  geom_hline(yintercept = -log10(threshold), linetype = "dashed") +
  # Colors (match your red theme)
  scale_color_manual(values = c("Both" = "orange", "LM DEG" = "red", "BM DEG" = "gold")) +
  labs(x = expression(-log[10](padj)~" LM DEG"),
    y = expression(-log[10](padj)~" BM DEG"),
    color = "Significance") +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  )


```


```{r Feature plot BM vs LM}
p <- FeaturePlot(onlyPT, features = c("BMScore1", "LMScore1"), min.cutoff = "q10", max.cutoff = "q90", blend = TRUE, cols = c("lightgrey", "yellow", "red"), combine = FALSE)

p[3] <- lapply(p[3], function(x) x+theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1)) + ylab("UMAP_1") + xlab("UMAP_2"))

p[4] <- lapply(p[4], function(x) x+theme(axis.line = element_line(color = "black"), axis.title = element_text(size = 15, family = "sans", color = "black"), axis.ticks = element_blank(), axis.text = element_blank(), panel.border = element_rect(color = "black", size = 1)) + ylab("LMScore") + xlab("BMScore"))

Fig6e <- p[3]
Fig6f <- p[4]

saveRDS(Fig6e, file = "Fig6e.rds")
saveRDS(Fig6f, file = "Fig6f.rds")

Fig6e
Fig6f
```
```{r FoldChange BMvsPT genes, only sign genes colored}
BMDeseqRes <- DeseqResBM[rownames(DeseqResBM) %in% rownames(DeseqRes), ]
LMDeseqRes <- DeseqRes[rownames(DeseqRes) %in% rownames(BMDeseqRes), ]

PTLMandBMFC <- data.frame(gene = rownames(LMDeseqRes), log2FoldChangeLM = LMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), LMDeseqRes$gene)], log2FoldChangeBM = BMDeseqRes$log2FoldChange[match(rownames(LMDeseqRes), BMDeseqRes$gene)], significance = ifelse(LMDeseqRes$gene %in% CommonUpGenes, "Up DEG Both", ifelse(LMDeseqRes$gene %in% CommonDownGenes, "Down DEG Both", ifelse(LMDeseqRes$gene %in% BMupGenes, "Up DEG BM", ifelse(LMDeseqRes$gene %in% upGenesDES, "Up DEG LM", ifelse(LMDeseqRes$gene %in% downGenesDES, "Down DEG LM", ifelse(LMDeseqRes$gene %in% BMdownGenes, "Down DEG BM", "Not significant")))))))

Vector <- c("Not significant", "Up DEG Both", "Down DEG Both", "Up DEG BM", "Up DEG LM", "Down DEG BM", "Down DEG LM")
PTLMandBMFC$significance <- factor(PTLMandBMFC$significance, levels = Vector)
PTLMandBMFC <- PTLMandBMFC[order(PTLMandBMFC$significance), ]

AllHighlightedGenes <- PTLMandBMFC[which(PTLMandBMFC$significance != "Not significant"), ]

Fig6d <- ggplot(PTLMandBMFC, aes(x = log2FoldChangeLM, y = log2FoldChangeBM, color = significance)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("Down DEG Both" = "#6BAED6", #"#00AFBB"
                                "Not significant" = "grey",
                                "Up DEG Both" = "orange",
                                "Up DEG BM" = "gold",
                                "Up DEG LM" = "red",
                                "Down DEG BM" = "#4e4e94",
                                "Down DEG LM" = "#088F8F")) +
  geom_vline(xintercept = c(-2, 2), linetype = "dashed") +
  geom_hline(yintercept = c(-2, 2), linetype = "dashed") +
  geom_point(size = 1) +
  geom_text_repel(data = AllHighlightedGenes, aes(label = gene), size = 2, max.overlaps = 20) +
  theme_minimal() +
  labs(title = "Fold change of LM and BM genes",
       x = "Average log2 Fold Change LM-PT",
       y = "Average log2 Fold Change BM-PT")

saveRDS(Fig6d, "Fig6d.rds")
Fig6d <- readRDS("Fig6d.rds")

Fig6d
```

###### Combining Plots
```{r Figure 1}
#(Fig1a | Fig1b | Fig1c) / Fig1d + plot_annotation(tag_levels = "A")
Fig1a
Fig1b
Fig1c
Fig1d
```

```{r Figure 2}
#(Fig2a / (Fig2b | Fig2c | Fig2d)) + plot_annotation(tag_levels = "A")
Fig2a
Fig2b
Fig2c
Fig2d
```

```{r Figure 3}
#((Fig3a / Fig3b) | Fig3c) / (Fig3d | Fig3e) + plot_annotation(tag_levels = "A")
Fig3a
Fig3b
Fig3c
Fig3d
Fig3e
```

```{r Figure 4}
#(Fig4a + plot_spacer() + plot_spacer()) / Fig4b + plot_annotation(tag_levels = "A")
Fig4a
Fig4b
Fig4c
```

```{r Figure 5}
#(Fig5a | Fig5b | Fig5c | Fig5d) / (Fig5e | Fig5f | plot_spacer()) + plot_annotation(tag_levels = "A")
Fig5a
Fig5b
Fig5c
Fig5d
Fig5e
Fig5f
```

```{r Figure 6}
Fig6a
Fig6b
Fig6c
Fig6d
```


