---
title: "OnlyLM"
execute: 
  message: false
  warning: false
format: 
  html:
    code-fold: true
    
---

```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
library(hdf5r)
library(SeuratData)
library(ggrepel)
library(tidyverse)

```
## Importing and orginising data

```{r}
#Setting working directory
setwd("~/T10")

#Importing data from files
#LM files
rawLM1 <- Read10X_h5(filename ="data/LM/PDX1/1_JEPDX1_6/sample_filtered_feature_bc_matrix.h5")
rawLM2 <- Read10X_h5(filename ="data/LM/PDX1/2_CHPDX1_5/sample_filtered_feature_bc_matrix.h5")
rawLM4 <- Read10X_h5(filename ="data/LM/PDX2/4_JEPDX2_5/sample_filtered_feature_bc_matrix.h5")
rawLM5 <- Read10X_h5(filename ="data/LM/PDX2/5_CHPDX2_5/sample_filtered_feature_bc_matrix.h5")
rawLM6 <- Read10X_h5(filename ="data/LM/PDX2/6_CHPDX2_7/sample_filtered_feature_bc_matrix.h5")
rawLM8 <- Read10X_h5(filename ="data/LM/PDX3/8_JEPDX3_5/sample_filtered_feature_bc_matrix.h5")
rawLM9 <- Read10X_h5(filename ="data/LM/PDX3/9_CHPDX3_3/sample_filtered_feature_bc_matrix.h5")
rawLM10 <- Read10X_h5(filename ="data/LM/PDX3/10_CHPDX3_5/sample_filtered_feature_bc_matrix.h5")

#Creating SEURAT objects
serLM1 <- CreateSeuratObject(rawLM1, project = "LM1")
serLM2 <- CreateSeuratObject(rawLM2, project = "LM2")
serLM4 <- CreateSeuratObject(rawLM4, project = "LM4")
serLM5 <- CreateSeuratObject(rawLM5, project = "LM5")
serLM6 <- CreateSeuratObject(rawLM6, project = "LM6")
serLM8 <- CreateSeuratObject(rawLM8, project = "LM8")
serLM9 <- CreateSeuratObject(rawLM9, project = "LM9")
serLM10 <- CreateSeuratObject(rawLM10, project = "LM10")

#Addning information layers
serLM1$PDX <- "LU-NB-1"
serLM2$PDX <- "LU-NB-1"
serLM4$PDX <- "LU-NB-2"
serLM5$PDX <- "LU-NB-2"
serLM6$PDX <- "LU-NB-2"
serLM8$PDX <- "LU-NB-3"
serLM9$PDX <- "LU-NB-3"
serLM10$PDX <- "LU-NB-3"

serLM1$type <- "LM"
serLM2$type <- "LM"
serLM4$type <- "LM"
serLM5$type <- "LM"
serLM6$type <- "LM"
serLM8$type <- "LM"
serLM9$type <- "LM"
serLM10$type <- "LM"

serLM1$mouse <- "mouse1"
serLM2$mouse <- "mouse2"
serLM4$mouse <- "mouse4"
serLM5$mouse <- "mouse5"
serLM6$mouse <- "mouse6"
serLM8$mouse <- "mouse8"
serLM9$mouse <- "mouse8"
serLM10$mouse <- "mouse10"

serLM1$batch <- "B1"
serLM2$batch <- "B1"
serLM4$batch <- "B1"
serLM5$batch <- "B1"
serLM6$batch <- "B2"
serLM8$batch <- "B2"
serLM9$batch <- "B2"
serLM10$batch <- "B2"

#Merging into 1 object
allData <- merge(serLM1, c(serLM2, serLM4, serLM5, serLM6, serLM8, serLM9, serLM10), add.cell.ids = c("LM1", "LM2", "LM4", "LM5", "LM6", "LM8", "LM9", "LM10"))
allData <- JoinLayers(allData)

#Remove unecessary objects
rm(serLM1, serLM2, serLM4, serLM5, serLM6, serLM8, serLM9, serLM10, rawLM1, rawLM2, rawLM4, rawLM5, rawLM6, rawLM8, rawLM9, rawLM10)
gc()

#Showing size of dataset
dim(allData)
table(allData$mouse)
table(allData$PDX)
```
Size of original dataset

##Filtering

```{r}
#Showing Mitochondrial, Ribosomal, Hemoglobin, Platelate markers
allData <- PercentageFeatureSet(allData, "^MT-", col.name = "percent_mito")
allData <- PercentageFeatureSet(allData, "^RP[SL]", col.name = "percent_ribo")
allData <- PercentageFeatureSet(allData, "^HB[^(P|E|S)]", col.name = "percent_hb")
allData <- PercentageFeatureSet(allData, "PECAM1|PF4", col.name = "percent_plat")

#Creating violin plots of QC variables
feats <- c("nFeature_RNA", "nCount_RNA", "percent_mito", "percent_ribo", "percent_hb", "percent_plat")
VlnPlot(allData, group.by = "mouse", split.by = "type", features = feats, pt.size = 0.05, ncol = 3)

```
We see high mitochondrial gene expression in many cells, this is indicative of cell membrane damage, allowing free RNA to leave the cytoplasm but not mitochondias due to their size. We want to filter out these cells. As well as cells with low total expression and low ribosomal expression. 

We can also limit the amount of genes we analyse by excluding genes expressed by only a few cells, as they will not hold any significant information for analysis. We will also exclude mitochondrial, ribosomal, and hemoglobin genes as physiological fluctuations in these may influence our analysis as well. 
```{r}
#Filter out cells and genes with low expression
selectCell <- WhichCells(allData, expression = nFeature_RNA > 100)
selectFeat <- rownames(allData)[Matrix::rowSums(allData[["RNA"]]$counts) > 10]
data.filt <- subset(allData, features = selectFeat, cells = selectCell)

#Filter out cells with high mitocondrial and low ribosomal expression
data.filt <- subset(data.filt, percent_mito < 20 & percent_ribo > 5)
```
Filtering done:
*Remove cells with less than 100 genes expressed.
*Remove genes expressed by less than 10 cells.
*Remove cells with above 20% of gene expression being mitochondrial genes.
*Remove cells with less than 5% of gene expression being ribosomal genes.
```{r}
#Showing size of dataset
dim(data.filt)
table(data.filt$mouse)
table(data.filt$PDX)


#New violin plot of QC variables
feats <- c("nFeature_RNA", "nCount_RNA", "percent_mito", "percent_ribo", "percent_hb")
VlnPlot(data.filt, group.by = "mouse", features = feats, pt.size = 0.05, ncol = 3) + NoLegend()
```

Size of dataset and visualization of quality control variables after first filtering. 

```{r}
# Filter Mitochondrial, Ribosomal and Hemoglobin genes
data.filt <- data.filt[!grepl("^MT-", rownames(data.filt)), ]
data.filt <- data.filt[ ! grepl("^RP[SL]", rownames(data.filt)), ]
data.filt <- data.filt[!grepl("^HB[^(P|E|S)]", rownames(data.filt)), ]
dim(data.filt)
```

Filtering done:
*Remove mitochondrial genes
*Remove ribosomal genes
*Remove Hemoglobin genes
Showing new dimensions fo dataset.

```{r}
suppressMessages(library(DoubletFinder))
run_doubletfinder_custom <- function(seu_sample_subset, multiplet_rate = NULL){
  # for debug
  #seu_sample_subset <- samp_split[[1]]
  # Print sample number
  print(paste0("Sample ", unique(seu_sample_subset[['batch']]), '...........')) 
  
  #Manually setting predicted doublet rate based on recovered cells per batch
  if(unique(seu_sample_subset[['batch']]) == 'B1'){
    multiplet_rate <- 0.008
  }
  if(unique(seu_sample_subset[['batch']]) == 'B2'){
    multiplet_rate <- 0.008
  }
  
  # Pre-process seurat object with standard seurat workflow --- 
  sample <- NormalizeData(seu_sample_subset, verbose = FALSE)
  sample <- FindVariableFeatures(sample, verbose = FALSE)
  sample <- ScaleData(sample, verbose = FALSE)
  sample <- RunPCA(sample, nfeatures.print = 10, verbose = FALSE)
  
  # Find significant PCs
  stdv <- sample[["pca"]]@stdev
  percent_stdv <- (stdv/sum(stdv)) * 100
  cumulative <- cumsum(percent_stdv)
  co1 <- which(cumulative > 90 & percent_stdv < 5)[1] 
  co2 <- sort(which((percent_stdv[1:length(percent_stdv) - 1] - 
                       percent_stdv[2:length(percent_stdv)]) > 0.1), 
              decreasing = T)[1] + 1
  min_pc <- min(co1, co2)
  
  # Finish pre-processing with min_pc
  sample <- RunUMAP(sample, dims = 1:min_pc, verbose = FALSE)
  sample <- FindNeighbors(object = sample, dims = 1:min_pc, verbose = FALSE)              
  sample <- FindClusters(object = sample, resolution = 0.1, verbose = FALSE)
  
  # pK identification (no ground-truth) 
  #introduces artificial doublets in varying props, merges with real data set and 
  # preprocesses the data + calculates the prop of artficial neighrest neighbours, 
  # provides a list of the proportion of artificial nearest neighbours for varying
  # combinations of the pN and pK
  sweep_list <- paramSweep(sample, PCs = 1:min_pc, sct = FALSE)   
  sweep_stats <- summarizeSweep(sweep_list)
  bcmvn <- find.pK(sweep_stats) # computes a metric to find the optimal pK value (max mean variance normalised by modality coefficient)
  # Optimal pK is the max of the bimodality coefficient (BCmvn) distribution
  optimal.pk <- bcmvn %>% 
    dplyr::filter(BCmetric == max(BCmetric)) %>%
    dplyr::select(pK)
  optimal.pk <- as.numeric(as.character(optimal.pk[[1]]))
  
  ## Homotypic doublet proportion estimate
  annotations <- sample@meta.data$seurat_clusters # use the clusters as the user-defined cell types
  homotypic.prop <- modelHomotypic(annotations) # get proportions of homotypic doublets
  
  #multiply by number of cells to get the number of expected multiplets
  nExp.poi <- round(multiplet_rate * nrow(sample@meta.data)) 
  nExp.poi.adj <- round(nExp.poi * (1 - homotypic.prop)) # expected number of doublets
  
  # run DoubletFinder
  sample <- doubletFinder(seu = sample, 
                          PCs = 1:min_pc, 
                          pK = optimal.pk, #the neighborhood size used to compute the number of artificial nearest neighbours
                          nExp = nExp.poi.adj) #number of expected real doublets
  
  # change name of metadata column with Singlet/Doublet information
  colnames(sample@meta.data)[grepl('DF.classifications.*', colnames(sample@meta.data))] <- "doublet_finder"
  
  #DimPlot(sample, group.by = "doublet_finder")
  
  #get the metadata column with singlet, doublet info
  double_finder_res <- sample@meta.data['doublet_finder'] 
  #add the cell IDs as new column to be able to merge correctly
  double_finder_res <- rownames_to_column(double_finder_res, "row_names") 
  return(double_finder_res)
}

data.split <- SplitObject(data.filt, split.by = "batch")
data.split <- lapply(data.split, run_doubletfinder_custom)

sglt_dblt_metadata <- data.frame(bind_rows(data.split)) # merge to a single dataframe
rownames(sglt_dblt_metadata) <- sglt_dblt_metadata$row_names # assign cell IDs to row names to ensure match
sglt_dblt_metadata$row_names <- NULL
head(sglt_dblt_metadata)
data.filt <- AddMetaData(data.filt, sglt_dblt_metadata, col.name = 'doublet_finder')

# Check how doublets singlets differ in QC measures per sample.
VlnPlot(data.filt, group.by = 'batch', split.by = "doublet_finder",
        features = c("nFeature_RNA", "nCount_RNA"), 
        ncol = 2, pt.size = 0) + theme(legend.position = 'right')

```
Plot showing higher counts and featured genes in doublets than singlets.

```{r}
suppressMessages(library(DoubletFinder))
data.filt <- data.filt[, data.filt@meta.data[, "doublet_finder"] == "Singlet"]
dim(data.filt)
table(data.filt$mouse)
table(data.filt$PDX)
```

Dataset size after filtering
Filtering done:
*Remove doublets. (amount of doublets filtered is based of predicted doublets, which increase with batch size. Liver metastasis sampels wer done with sample 1-4 in one batch and 5-8 in one. All primary tumor samples were run in individual, larger batches, leading to a higher frequency of doublets as well.)

##Preparation of data, Integration
Normalizing data: Each cell has a different total numbers of reads. Normalising compensates for that.  
Finding variable features: Finding genes with high variance.
Scaling data: Centers and equalizes the value of each genes. Sets mean = 0 and variance = 1.
PCA: Weights the found variable features to capture as muchect.

```{r}
data.filt[["RNA"]] <- split(data.filt[["RNA"]], f = data.filt$batch)

data.filt <- NormalizeData(data.filt, verbose = FALSE)
data.filt <- FindVariableFeatures(data.filt, verbose = FALSE)
data.filt <- ScaleData(data.filt, verbose = FALSE)
data.pca <- RunPCA(data.filt, verbose = FALSE)

#Integrating
data.integ <- IntegrateLayers(object = data.pca, method = HarmonyIntegration, orig.reduction = "pca", new.reduction = "harmony",
                             verbose = FALSE)
data.integ[["RNA"]] <- JoinLayers(data.integ[["RNA"]])


#Printing components of PCA
wrap_plots(
  DimPlot(data.pca, reduction = "pca", group.by = "PDX", dims = 1:2),
  DimPlot(data.pca, reduction = "pca", group.by = "PDX", dims = 3:4),
  DimPlot(data.pca, reduction = "pca", group.by = "PDX", dims = 5:6),
  ncol = 3
) + plot_layout(guides = "collect")

```
Showing the variance captured by first 6 PCA dimensions. (Maybe not relevant to show?) 
```{r}
#What genes contribute to PCs
VizDimLoadings(data.pca, dims = 1:4, reduction = "pca", ncol = 4, balanced = T)
```
Showing what genes contributed first 4 PCA dimensions. (Relevant?)

##Visualising using UMAP and tSNE
UMAP is better for visualizing variability between cell clusters. tSNE is better for visualizing in-group variability. 

```{r}
#UMAP on integrated data
data.umap <- RunUMAP(data.integ, dims = 1:30, reduction = "harmony", reduction.name = "umap_harmony", verbose = FALSE)
#UMAP on NonIntegrated data
data.umapNonInt <- RunUMAP(data.pca, dims = 1:30, reduction = "pca", verbose = FALSE)

#tSNE on integrated data
data.tsne <- RunTSNE(
  data.integ,
  reduction = "harmony", dims = 1:15,
  perplexity = 30,
  max_iter = 1000,
  theta = 0.5,
  eta = 200,
  num_threads = 0,
  verbose = FALSE
)
#tSNE on NonIntegrated data
data.tsneNonInt <- RunTSNE(
  data.pca,
  reduction = "pca", dims = 1:15,
  perplexity = 30,
  max_iter = 1000,
  theta = 0.5,
  eta = 200,
  num_threads = 0,
  vebrose = FALSE
)

#Summorize based on mouse
wrap_plots(
  DimPlot(data.umapNonInt, reduction = "umap", group.by = "PDX") + ggtitle("NonIntegrated UMAP"),
  DimPlot(data.umap, reduction = "umap_harmony", group.by = "PDX") + ggtitle("Integrated UMAP"),
  DimPlot(data.tsneNonInt, reduction = "tsne", group.by = "PDX") + ggtitle("NonIntegrated tSNE"),
  DimPlot(data.tsne, reduction = "tsne", group.by = "PDX") + ggtitle("Integrated tSNE"),
  ncol = 2
) + plot_layout(guides = "collect")
```
Showing UMAP and tSNE based on mouse and before and after compensating for batch effect(integrating)

##Clustering
```{r}
#Clustering
data.umap <- FindNeighbors(data.umap, dims = 1:20, verbose = FALSE)
data.umap <- FindClusters(data.umap, resolution = 0.5, verbose = FALSE)

data.tsne <- FindNeighbors(data.tsne, dims = 1:20, verbose = FALSE)
data.tsne <- FindClusters(data.tsne, resolution = 0.5, verbose = FALSE)

#Visualising if calculated clusters correspond to celltypes
wrap_plots(
  DimPlot(data.umap, reduction = "umap_harmony") + ggtitle("UMAP clustered"),
  DimPlot(data.umap, reduction = "umap_harmony", group.by = "PDX") + ggtitle("UMAP PDX"),
  DimPlot(data.umap, reduction = "umap_harmony", group.by = "mouse") + ggtitle("UMAP mouse"),
  DimPlot(data.tsne, reduction = "tsne") + ggtitle("tSNE clustered"),
  DimPlot(data.tsne, reduction = "tsne", group.by = "PDX") + ggtitle("tSNE PDX"),
  DimPlot(data.tsne, reduction = "tsne", group.by = "mouse") + ggtitle("tSNE mouse"),
  ncol = 3
) + plot_layout(guides = "collect")
```

Visualizing if calculated clusters and how they correspond to PDX and mouse
